## Binary Tree Inorder Traversal {#lc-inorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 94
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-inorder-traversal/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Return the inorder traversal (left-root-right) of a binary tree.

### Solution - Iterative Stack

```
public List<Integer> inorderTraversal(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    Deque<TreeNode> stack = new ArrayDeque<>();
    TreeNode current = root;

    while (current != null || !stack.isEmpty()) {
        while (current != null) {
            stack.push(current);
            current = current.left;
        }
        current = stack.pop();
        result.add(current.val);
        current = current.right;
    }
    return result;
}
```

### Solution - Recursive

```
public List<Integer> inorderTraversalRecursive(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    dfs(root, result);
    return result;
}

private void dfs(TreeNode node, List<Integer> lc-list) {
    if (node == null) {
        return;
    }
    dfs(node.left, lc-list);
    lc-list.add(node.val);
    dfs(node.right, lc-list);
}
```


## Validate Binary Search Tree {#lc-validate-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 98
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/validate-binary-search-tree/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), [Recursion](glossary-of-algorithm.html#recursion)

### Description

Check whether a binary tree is a valid BST.

### Solution - Inorder Bounds

Use DFS while carrying min/max bounds.

```
public boolean isValidBST(TreeNode root) {
    return validate(root, null, null);
}

private boolean validate(TreeNode node, Integer low, Integer high) {
    if (node == null) {
        return true;
    }
    if ((low != null && node.val <= low) || (high != null && node.val >= high)) {
        return false;
    }
    return validate(node.left, low, node.val) && validate(node.right, node.val, high);
}
```


## Binary Tree Zigzag Level Order Traversal {#lc-zigzag-level-order}

```
### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 103
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-zigzag-level-order-traversal/
- **Tags**: `Tree`, `BFS`
- **Techniques**: [Breadth First Search](glossary-of-algorithm.html#bfs)

### Description

Return the zigzag level order traversal of a binary tree (left-to-right, then right-to-left, alternating each level).

### Examples

```
Input: [3,9,20,null,null,15,7]
Output: [[3],[20,9],[15,7]]
```

### Constraints

- `0 <= nodes <= 2000`

### Solution - BFS with Direction Flag

#### Walkthrough

Perform level-order traversal. Track a boolean indicating whether to append values normally or reversed. For every level, either append values or insert at the front, and flip the flag.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Implementation Steps

1. Initialize queue with root and a boolean `leftToRight = true`.
2. For each level, collect values in a lc-list.
3. If `leftToRight`, append; otherwise insert at front or use reverse lc-list.
4. Toggle `leftToRight` each level.

#### Java Code
```java
public List<List<Integer>> zigzagLevelOrder(TreeNode root) {
    List<List<Integer>> result = new ArrayList<>();
    if (root == null) {
        return result;
    }

    Queue<TreeNode> queue = new LinkedList<>();
    queue.offer(root);
    boolean leftToRight = true;

    while (!queue.isEmpty()) {
        int size = queue.size();
        LinkedList<Integer> level = new LinkedList<>();

        for (int i = 0; i < size; i++) {
            TreeNode node = queue.poll();
            if (leftToRight) {
                level.addLast(node.val);
            } else {
                level.addFirst(node.val);
            }

            if (node.left != null) {
                queue.offer(node.left);
            }
            if (node.right != null) {
                queue.offer(node.right);
            }
        }

        result.add(level);
        leftToRight = !leftToRight;
    }

    return result;
}
```


## Maximum Depth of Binary Tree {#lc-max-depth}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 104
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/maximum-depth-of-binary-tree/
- **Tags**: `Tree`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), DFS

### Description

Return the maximum depth (number of nodes along the longest path from root down to a leaf).

### Solution - DFS

```
public int maxDepth(TreeNode root) {
    if (root == null) {
        return 0;
    }
    return 1 + Math.max(maxDepth(root.left), maxDepth(root.right));
}
```


## Construct Binary Tree from Preorder and Inorder Traversal {#lc-build-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 105
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/construct-binary-tree-from-preorder-and-inorder-traversal/
- **Tags**: `Tree`
- **Techniques**: [Divide and Conquer](glossary-of-algorithm.html#dnc), [Recursion](glossary-of-algorithm.html#recursion), [Hash Table](glossary-of-algorithm.html#hash-table)

### Description

Build a binary tree from preorder and inorder traversals.

### Solution - Hash Map Recursion

```
public TreeNode buildTree(int[] preorder, int[] inorder) {
    Map<Integer, Integer> indexMap = new HashMap<>();
    for (int i = 0; i < inorder.length; i++) {
        indexMap.put(inorder[i], i);
    }
    return build(preorder, 0, preorder.length - 1, 0, inorder.length - 1, indexMap);
}

private TreeNode build(int[] preorder, int preL, int preR, int inL, int inR, Map<Integer, Integer> indexMap) {
    if (preL > preR) {
        return null;
    }
    int rootVal = preorder[preL];
    TreeNode root = new TreeNode(rootVal);
    int inorderRootIndex = indexMap.get(rootVal);
    int leftSize = inorderRootIndex - inL;
    root.left = build(preorder, preL + 1, preL + leftSize, inL, inorderRootIndex - 1, indexMap);
    root.right = build(preorder, preL + leftSize + 1, preR, inorderRootIndex + 1, inR, indexMap);
    return root;
}
```

## Binary Tree Level Order Traversal II {#lc-level-order-bottom}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 107
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-level-order-traversal-ii/
- **Tags**: `Tree`, `BFS`
- **Techniques**: [Breadth First Search](glossary-of-algorithm.html#bfs)

### Description

Return the level order traversal of a binary tree from the bottom level up to the root.

### Examples

```
Input: [3,9,20,null,null,15,7]
Output: [[15,7],[9,20],[3]]
```

### Constraints

- `0 <= nodes <= 2000`

### Solution - Reverse BFS

#### Walkthrough

Perform BFS collecting nodes level by level. Instead of appending to the end, insert each level at the front of a linked lc-list (or append and reverse at the end).

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Implementation Steps

1. Run BFS using a queue.
2. For each level size, dequeue nodes, record values, enqueue children.
3. Insert level lc-list at the start of the result.

#### Java Code
```java
public List<List<Integer>> levelOrderBottom(TreeNode root) {
    LinkedList<List<Integer>> result = new LinkedList<>();
    if (root == null) {
        return result;
    }

    Queue<TreeNode> queue = new LinkedList<>();
    queue.offer(root);

    while (!queue.isEmpty()) {
        int size = queue.size();
        List<Integer> level = new ArrayList<>(size);

        for (int i = 0; i < size; i++) {
            TreeNode node = queue.poll();
            level.add(node.val);
            if (node.left != null) {
                queue.offer(node.left);
            }
            if (node.right != null) {
                queue.offer(node.right);
            }
        }

        result.addFirst(level);
    }

    return result;
}
```


## Convert Sorted Array to BST {#lc-sorted-array-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 108
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/convert-sorted-array-to-binary-search-tree/
- **Tags**: `Tree`, `Divide and Conquer`
- **Techniques**: [Divide and Conquer](glossary-of-algorithm.html#dnc), [Recursion](glossary-of-algorithm.html#recursion)

### Description

Convert a sorted array to a height-balanced BST.

### Solution - Divide and Conquer

```
public TreeNode sortedArrayToBST(int[] nums) {
    return build(nums, 0, nums.length - 1);
}

private TreeNode build(int[] nums, int left, int right) {
    if (left > right) {
        return null;
    }
    int mid = left + (right - left) / 2;
    TreeNode node = new TreeNode(nums[mid]);
    node.left = build(nums, left, mid - 1);
    node.right = build(nums, mid + 1, right);
    return node;
}
```


## Balanced Binary Tree {#lc-balanced-binary-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 110
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/lc-balanced-binary-tree/
- **Tags**: `Tree`, `Binary Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), [Recursion](glossary-of-algorithm.html#recursion)

### Description

Given a binary tree, determine if it is **height-balanced**.

For this problem, a height-balanced binary tree is defined as:

> A binary tree in which the depth of the two subtrees of **every** node never differ by more than 1.

### Examples

**Example 1:**
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Balanced tree example', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[
level 1/.style={sibling distance=10em},
level 2/.style={sibling distance=5em},
level 3/.style={sibling distance=5em},
level distance=2.5em,
]
\node {3}
    child { node {9}}
    child { node {20}
        child { node {15} }
        child { node {7} }
    }
;
\end{tikzpicture}
```
```
Input: root = [3,9,20,null,null,15,7]
Output: true
Explanation: Height of left subtree = 1, height of right subtree = 2, difference = 1 (balanced)
```

**Example 2:**
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Unbalanced tree example', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[
    level 1/.style={sibling distance=10em},
    level 2/.style={sibling distance=5em},
    level 3/.style={sibling distance=5em},
    level distance=2.5em,
    ]
\node {1}
    child { node {2}
        child {  node {3}
            child {  node {4}}
            child {  node {4}}
        }
        child {  node {3}}
    }
    child { node {2}}
;
\end{tikzpicture}
```
```
Input: root = [1,2,2,3,3,null,null,4,4]
Output: false
Explanation: Left subtree has much greater depth than right subtree
```

**Example 3:**
```
Input: root = []
Output: true
```

### Constraints

- The number of nodes in the tree is in the range `[0, 5000]`
- `-10^4 <= Node.val <= 10^4`

### Solution - Recursive DFS

#### Walkthrough

Use a recursive depth-first search to compute the height of each subtree.

For each node:
1. Recursively compute the height of the left subtree
2. Recursively compute the height of the right subtree
3. Check if the absolute difference between heights is greater than 1
4. If unbalanced at any node, propagate `-1` upward to indicate imbalance
5. Otherwise, return the maximum height + 1

The key insight is to combine height calculation with balance checking in a single pass.

#### Analysis

- **Time Complexity**: O(n) - Each node is visited exactly once
- **Space Complexity**: O(h) - Recursion call stack depth equals tree height h, where h can be log(n) for balanced trees or n for skewed trees

#### Implementation Steps


1. If root is null, return 0 (base case)
2. Recursively get height of left subtree: `lHeight = getHeight(root.left)`
3. Recursively get height of right subtree: `rHeight = getHeight(root.right)`
4. If either subtree returned -1 (unbalanced), propagate -1 upward
5. If `|lHeight - rHeight| > 1`, return -1 (unbalanced)
6. Otherwise, return `max(lHeight, rHeight) + 1`

#### Java Code

```java
public boolean isBalanced(TreeNode root) {
    return getHeight(root) != -1;
}

private int getHeight(TreeNode root) {
    if (root == null) {
        return 0;
    }

    int lHeight = getHeight(root.left);
    int rHeight = getHeight(root.right);

    // If left or right subtree is unbalanced, propagate -1
    if (lHeight == -1 || rHeight == -1) {
        return -1;
    }

    // If current node is unbalanced, return -1
    if (Math.abs(lHeight - rHeight) > 1) {
        return -1;
    }

    // Return height of current subtree
    return Math.max(lHeight, rHeight) + 1;
}
```











### Solution - BFS with One Loops
#### Walkthrough
Create a queue and push node of the same level in a two dimensional array lc-list while traversing the tree with
BFS algorithm.

#### Analysis
Breadth first traversal has time complexity of O(n), as every node is visited and thus  the space
complexity is also O(n).

#### Implementation Steps

 bfs \@ref(bfs)

#### Java Code - BFS with One Loop
```java
public List<List<Integer>> levelOrder(TreeNode root) {
    List<List<Integer>> levels = new ArrayList<>();
    if (root == null) {
        return levels;
    }
    Queue<TreeNode> currentLevelQueue = new LinkedList<TreeNode>();
    Queue<TreeNode> nextLevelQueue = new LinkedList<TreeNode>();

    currentLevelQueue.offer(root);
    //init level to null
    List<Integer> level = null;

    while(!currentLevelQueue.isEmpty()) {
        TreeNode node = currentLevelQueue.poll();

        //starting a new level
        if (level == null) {
            level = new LinkedList<Integer>();
            levels.add(level);
        }
        level.add(node.val);

        //push children to next level queue
        if (node.left != null) {
            nextLevelQueue.offer(node.left);
        }
        if (node.right != null) {
            nextLevelQueue.offer(node.right);
        }

        //swap queue and next, where next collects all nodes for next level
        if (currentLevelQueue.isEmpty()) {
            Queue<TreeNode> temp = currentLevelQueue;
            currentLevelQueue = nextLevelQueue;
            nextLevelQueue = temp;
            //init level to null
            level = null;
        }
    }
    return levels;
}
```

### Solution - DFS
#### Walkthrough
We could traverse the tree recursively with \textbf{PreOrder} strategy, and additionally pass current level
into recursive method.

#### Analysis
DFS has time complexity of O(n) as every node is visited and thus Auxiliary Space is also O(n).

#### Implementation Steps

 dfs \@ref(dfs)

#### Java Code - DFS
```java
public List<List<Integer>> levelOrder(TreeNode root) {
    List<List<Integer>> ans = new ArrayList<>();
    traverse(root, 0, ans);

    return ans;
}

private void traverse(TreeNode root, int level, List<List<Integer>> ans) {
    if(root == null){
        return;
    }

    if(ans.size()<=level){
        ans.add(new ArrayList<>());
    }

    ans.get(level).add(root.val);

    traverse(root.left, level + 1, ans);
    traverse(root.right, level + 1, ans);
}
```


## Minimum Depth of Binary Tree {#lc-min-depth}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 111
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/minimum-depth-of-binary-tree/
- **Tags**: `Tree`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), DFS

### Description

Return the minimum depth from root to the nearest leaf.

### Solution - DFS

```
public int minDepth(TreeNode root) {
    if (root == null) {
        return 0;
    }
    if (root.left == null) {
        return 1 + minDepth(root.right);
    }
    if (root.right == null) {
        return 1 + minDepth(root.left);
    }
    return 1 + Math.min(minDepth(root.left), minDepth(root.right));
}
```


## Path Sum {#lc-path-sum}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 112
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/lc-path-sum/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Determine if a binary tree has a root-to-leaf path whose node values sum to a target.

### Examples

```
Input: root = [5,4,8,11,null,13,4,7,2,null,null,null,1], target = 22
Output: true
```

### Constraints

- `0 <= nodes <= 5000`

### Solution - DFS with Running Sum

#### Walkthrough

Traverse the tree, accumulating the current sum along the path. When a leaf is reached, check if the sum equals the target. Short-circuit as soon as a valid path is found.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Java Code
```
public boolean hasPathSum(TreeNode root, int targetSum) {
    if (root == null) {
        return false;
    }
    if (root.left == null && root.right == null) {
        return targetSum == root.val;
    }
    int remaining = targetSum - root.val;
    return hasPathSum(root.left, remaining) || hasPathSum(root.right, remaining);
}
```


## Path Sum II {#lc-path-sum-ii}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 113
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lc-path-sum-ii/
- **Tags**: `Tree`, `Backtracking`
- **Techniques**: [Backtracking: Choose-Explore-Unchoose](#backtracking-choose-explore-unchoose-pattern)

### Description

Find all root-to-leaf paths where the sum of node values equals a target.

### Solution - Backtracking

#### Walkthrough

Traverse the tree, maintaining the current path lc-list and remaining sum. When a leaf reaches zero remaining sum, append a copy of the path to the answer. Backtrack after exploring each branch.

#### Analysis

- **Time Complexity**: O(n * h)
- **Space Complexity**: O(h)

#### Java Code
```
public List<List<Integer>> pathSum(TreeNode root, int targetSum) {
    List<List<Integer>> result = new ArrayList<>();
    backtrack(root, targetSum, new ArrayList<>(), result);
    return result;
}

private void backtrack(TreeNode node, int remain, List<Integer> path, List<List<Integer>> result) {
    if (node == null) {
        return;
    }

    path.add(node.val);
    remain -= node.val;

    if (node.left == null && node.right == null && remain == 0) {
        result.add(new ArrayList<>(path));
    } else {
        backtrack(node.left, remain, path, result);
        backtrack(node.right, remain, path, result);
    }

    path.remove(path.size() - 1);
}
```


## Populating Next Right Pointers in Each Node {#lc-connect-next}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 116
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/populating-next-right-pointers-in-each-node/
- **Tags**: `Tree`, `Breadth First Search`
- **Techniques**: [Breadth First Search](glossary-of-algorithm.html#bfs)

### Description

Connect each node's `next` pointer to its right neighbor on the same level in a perfect binary tree.

### Solution - O(1) Level Traversal

```
public Node connect(Node root) {
    if (root == null) {
        return null;
    }
    Node leftmost = root;
    while (leftmost.left != null) {
        Node head = leftmost;
        while (head != null) {
            head.left.next = head.right;
            if (head.next != null) {
                head.right.next = head.next.left;
            }
            head = head.next;
        }
        leftmost = leftmost.left;
    }
    return root;
}
```


## Binary Tree Preorder Traversal {#lc-preorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 144
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-preorder-traversal/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Return the preorder traversal (root-left-right) of a binary tree.

### Solution - Iterative Stack

#### Walkthrough

Push root to a stack. Pop, record value, then push right and left children (right first so left is processed next). Repeat until stack empty.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Java Code
```
public List<Integer> preorderTraversal(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    if (root == null) {
        return result;
    }

    Deque<TreeNode> stack = new ArrayDeque<>();
    stack.push(root);

    while (!stack.isEmpty()) {
        TreeNode node = stack.pop();
        result.add(node.val);
        if (node.right != null) {
            stack.push(node.right);
        }
        if (node.left != null) {
            stack.push(node.left);
        }
    }
    return result;
}
```

### Solution - Recursive

```
public List<Integer> preorderTraversalRecursive(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    dfs(root, result);
    return result;
}

private void dfs(TreeNode node, List<Integer> lc-list) {
    if (node == null) {
        return;
    }
    lc-list.add(node.val);
    dfs(node.left, lc-list);
    dfs(node.right, lc-list);
}
```


## Binary Tree Postorder Traversal {#lc-postorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 145
- **Difficulty**: Hard
- **URL**: https://leetcode.com/problems/binary-tree-postorder-traversal/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Return the postorder traversal (left-right-root) of a binary tree.

### Solution - Iterative Stack

```
public List<Integer> postorderTraversal(TreeNode root) {
    LinkedList<Integer> result = new LinkedList<>();
    if (root == null) {
        return result;
    }

    Deque<TreeNode> stack = new ArrayDeque<>();
    stack.push(root);
    while (!stack.isEmpty()) {
        TreeNode node = stack.pop();
        result.addFirst(node.val);
        if (node.left != null) {
            stack.push(node.left);
        }
        if (node.right != null) {
            stack.push(node.right);
        }
    }
    return result;
}
```

### Solution - Recursive

```
private void dfs(TreeNode node, List<Integer> lc-list) {
    if (node == null) {
        return;
    }
    dfs(node.left, lc-list);
    dfs(node.right, lc-list);
    lc-list.add(node.val);
}
```


## Binary Tree Upside Down {#lc-upside-down}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 156
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-upside-down/
- **Tags**: `Tree`
- **Techniques**: [Recursion](glossary-of-algorithm.html#recursion)

### Description

Flip the tree so that every right child becomes a left sibling.

### Solution - Recursive Rotation

```
public TreeNode upsideDownBinaryTree(TreeNode root) {
    if (root == null || root.left == null) {
        return root;
    }
    TreeNode newRoot = upsideDownBinaryTree(root.left);
    root.left.left = root.right;
    root.left.right = root;
    root.left = null;
    root.right = null;
    return newRoot;
}
```


## Count Univalue Subtrees {#lc-count-univalue}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 250
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/count-univalue-subtrees/
- **Tags**: `Tree`, `DFS`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), [Recursion](glossary-of-algorithm.html#recursion)

### Description

Count subtrees where every node has the same value.

### Solution - DFS

Traverse bottom-up. A node is univalue if both children are univalue and match the current value.

```
public int countUnivalSubtrees(TreeNode root) {
    int[] count = new int[1];
    dfs(root, count);
    return count[0];
}

private boolean dfs(TreeNode node, int[] count) {
    if (node == null) {
        return true;
    }
    boolean left = dfs(node.left, count);
    boolean right = dfs(node.right, count);
    if (!left || !right) {
        return false;
    }
    if (node.left != null && node.left.val != node.val) {
        return false;
    }
    if (node.right != null && node.right.val != node.val) {
        return false;
    }
    count[0]++;
    return true;
}
```


## Inorder Successor in BST {#lc-inorder-successor}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 285
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/inorder-successor-in-bst/
- **Tags**: `Tree`
- **Techniques**: [Binary Search Tree](glossary-of-algorithm.html#bst)

### Description

Find the inorder successor of a given node in a BST.

### Solution - BST Properties

```
public TreeNode inorderSuccessor(TreeNode root, TreeNode p) {
    TreeNode succ = null;
    while (root != null) {
        if (p.val < root.val) {
            succ = root;
            root = root.left;
        } else {
            root = root.right;
        }
    }
    return succ;
}
```


## Binary Tree Longest Consecutive Sequence {#lc-longest-consecutive}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 298
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-longest-consecutive-sequence/
- **Tags**: `Tree`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Find the length of the longest path where consecutive nodes differ by +1.

### Solution - DFS Tracking

```
public int longestConsecutive(TreeNode root) {
    return dfs(root, null, 0);
}

private int dfs(TreeNode node, TreeNode parent, int length) {
    if (node == null) {
        return length;
    }
    if (parent != null && node.val == parent.val + 1) {
        length++;
    } else {
        length = 1;
    }
    int left = dfs(node.left, node, length);
    int right = dfs(node.right, node, length);
    return Math.max(length, Math.max(left, right));
}
```


## Find Leaves of Binary Tree {#lc-find-leaves}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 366
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/find-leaves-of-binary-tree/
- **Tags**: `Tree`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs), [Recursion](glossary-of-algorithm.html#recursion)

### Description

Collect nodes layer by layer from leaves upward.

### Solution - Height DFS

Compute the height of each node (distance from leaf). Use height as index to append values to the result lc-list.

```
public List<List<Integer>> findLeaves(TreeNode root) {
    List<List<Integer>> result = new ArrayList<>();
    height(root, result);
    return result;
}

private int height(TreeNode node, List<List<Integer>> result) {
    if (node == null) {
        return -1;
    }
    int h = Math.max(height(node.left, result), height(node.right, result)) + 1;
    if (result.size() == h) {
        result.add(new ArrayList<>());
    }
    result.get(h).add(node.val);
    return h;
}
```


## Path Sum III {#lc-path-sum-iii}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 437
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lc-path-sum-iii/
- **Tags**: `Tree`, `Prefix Sum`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Count the number of paths in a binary tree that sum to a target. Paths may start and end anywhere but must go downward.

### Solution - Prefix Sum DFS

#### Walkthrough

Use DFS while tracking a running prefix sum and a hash map of prefix counts. For each node, check how many previous prefix sums equal `currentSum - target`. Update the map while exploring children and backtrack counts afterward.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Java Code
```
public int pathSum(TreeNode root, int targetSum) {
    Map<Integer, Integer> prefix = new HashMap<>();
    prefix.put(0, 1);
    return dfs(root, 0, targetSum, prefix);
}

private int dfs(TreeNode node, int current, int target, Map<Integer, Integer> prefix) {
    if (node == null) {
        return 0;
    }

    current += node.val;
    int count = prefix.getOrDefault(current - target, 0);
    prefix.put(current, prefix.getOrDefault(current, 0) + 1);

    count += dfs(node.left, current, target, prefix);
    count += dfs(node.right, current, target, prefix);

    prefix.put(current, prefix.get(current) - 1);
    return count;
}
```


## Diameter of Binary Tree {#lc-diameter}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 543
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/diameter-of-binary-tree/
- **Tags**: `Tree`
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#dfs)

### Description

Return the length of the longest path between any two nodes.

### Solution - Depth Helper

Use DFS to compute depth of each subtree and track the maximum `leftDepth + rightDepth`.

```
public int diameterOfBinaryTree(TreeNode root) {
    int[] diameter = new int[1];
    depth(root, diameter);
    return diameter[0];
}

private int depth(TreeNode node, int[] diameter) {
    if (node == null) {
        return 0;
    }
    int left = depth(node.left, diameter);
    int right = depth(node.right, diameter);
    diameter[0] = Math.max(diameter[0], left + right);
    return Math.max(left, right) + 1;
}
```
