## Binary Tree Inorder Traversal {#lc-inorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 94
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/binary-tree-inorder-traversal/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), [In-Order Traversal](#intro-tree-inorder), [Stack](glossary-of-algorithm.html#glossary-stack), Tree

### Description

Return the inorder traversal (left-root-right) of a binary tree.

### Examples

**Example 1: Regular Binary Tree**
```
Input:
    1
   / \
  2   3
Output: [2,1,3]
Explanation: In-order visits left subtree first (2), then root (1), then right subtree (3)
```

**Example 2: Empty Tree**
```
Input: null
Output: []
```

**Example 3: Skewed Tree**
```
Input:
    1
     \
      2
       \
        3
Output: [1,2,3]
Explanation: No left children, visits nodes in order down the right spine
```

### Constraints

- The number of nodes in the tree is in the range `[0, 100]`
- `-100 <= Node.val <= 100`

### Solution - Iterative Stack

#### Walkthrough

Use a stack to simulate recursion and always dive left first:

1. Maintain `current` pointer starting at root.
2. **Traverse left spine**: push nodes while moving `current = current.left` until null.
3. Pop from stack, record value (the "root" between left and right), then shift to right subtree `current = node.right`.
4. Repeat until both `current` is null and stack is empty.

This yields nodes in left-root-right order without recursion.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n) worst case stack on a skewed tree


#### Code - Java

```java
public List<Integer> inorderTraversal(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    Deque<TreeNode> stack = new ArrayDeque<>();
    TreeNode current = root;

    while (current != null || !stack.isEmpty()) {
        while (current != null) {
            stack.push(current);
            current = current.left;
        }
        current = stack.pop();
        result.add(current.val);
        current = current.right;
    }
    return result;
}
```

### Solution - Recursive

#### Walkthrough

Recursively traverse left subtree, visit the current node, then traverse right subtree. This mirrors the inorder definition and appends values in left-root-right order.

The recursive approach is more intuitive and directly reflects the definition of in-order traversal:
1. **Base case**: If node is null, return immediately
2. **Recurse left**: Visit all nodes in the left subtree
3. **Process current**: Add current node's value to result
4. **Recurse right**: Visit all nodes in the right subtree

#### Analysis

- **Time Complexity**: O(n) - Each node is visited exactly once
- **Space Complexity**: O(h) - Recursion call stack depth where h is tree height (O(n) worst case for skewed tree)

#### Code - Java

```java
public List<Integer> inorderTraversalRecursive(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    dfs(root, result);
    return result;
}

private void dfs(TreeNode node, List<Integer> list) {
    if (node == null) {
        return;
    }
    dfs(node.left, list);
    list.add(node.val);
    dfs(node.right, list);
}
```


## Validate Binary Search Tree {#lc-validate-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 98
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/validate-binary-search-tree/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Check whether a binary tree is a valid BST.

### Examples

**Example 1: Valid BST**
```
Input:
    2
   / \
  1   3
Output: true
Explanation: All left values (1) are less than root (2), all right values (3) are greater
```

**Example 2: Invalid BST**
```
Input:
    5
   / \
  1   4
     / \
    3   6
Output: false
Explanation: Node 4's left child (3) violates the BST property for subtree rooted at 5
```

**Example 3: Single Node**
```
Input: 1
Output: true
Explanation: A single node is always a valid BST
```

### Constraints

- The number of nodes in the tree is in the range `[1, 10^4]`
- `-2^31 <= Node.val <= 2^31 - 1`

### Solution - Inorder Bounds

#### Walkthrough

Perform DFS carrying the valid range for each subtree:

1. Start with no bounds at root.
2. For a node, if its value is not strictly within `(low, high)`, return false.
3. Recurse left with updated `high = node.val` (all left nodes must be smaller).
4. Recurse right with updated `low = node.val` (all right nodes must be larger).
5. Return true only if both subtrees are valid.

Passing bounds prevents false positives that would slip through a simple local child comparison.

#### Analysis

- **Time Complexity**: O(n) — visit each node once
- **Space Complexity**: O(h) recursion stack (h = tree height)

#### Code - Java

```
public boolean isValidBST(TreeNode root) {
    return validate(root, null, null);
}

private boolean validate(TreeNode node, Integer low, Integer high) {
    if (node == null) {
        return true;
    }
    if ((low != null && node.val <= low) || (high != null && node.val >= high)) {
        return false;
    }
    return validate(node.left, low, node.val) && validate(node.right, node.val, high);
}
```

#### Code - Golang

```go
func isValidBST(root *TreeNode) bool {
    return validate(root, math.MinInt32 - 1, math.MaxInt32 + 1)
}

func validate(node *TreeNode, low int, high int) bool {
    if node == nil {
        return true;
    }

    if node.Val <= low || node.Val >= high {
        return false;
    }

    return validate(node.Left, low, node.Val) && validate(node.Right, node.Val, high);
}
```

## Binary Tree Zigzag Level Order Traversal {#lc-zigzag-level-order}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 103
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-zigzag-level-order-traversal/
- **Tags**:
- **Techniques**: [Breadth First Search](#intro-tree-bfs), Tree

### Description

Return the zigzag level order traversal of a binary tree (left-to-right, then right-to-left, alternating each level).

### Examples

```
Input: [3,9,20,null,null,15,7]
Output: [[3],[20,9],[15,7]]
```

### Constraints

- `0 <= nodes <= 2000`

### Solution - BFS with Direction Flag

#### Walkthrough

Perform level-order traversal. Track a boolean indicating whether to append values normally or reversed. For every level, either append values or insert at the front, and flip the flag.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Implementation Steps

1. Initialize queue with root and a boolean `leftToRight = true`.
2. For each level, collect values in a list.
3. If `leftToRight`, append; otherwise insert at front or use reverse list.
4. Toggle `leftToRight` each level.

#### Code - Java
```java
public List<List<Integer>> zigzagLevelOrder(TreeNode root) {
    List<List<Integer>> result = new ArrayList<>();
    if (root == null) {
        return result;
    }

    Queue<TreeNode> queue = new LinkedList<>();
    queue.offer(root);
    boolean leftToRight = true;

    while (!queue.isEmpty()) {
        int size = queue.size();
        LinkedList<Integer> level = new LinkedList<>();

        for (int i = 0; i < size; i++) {
            TreeNode node = queue.poll();
            if (leftToRight) {
                level.addLast(node.val);
            } else {
                level.addFirst(node.val);
            }

            if (node.left != null) {
                queue.offer(node.left);
            }
            if (node.right != null) {
                queue.offer(node.right);
            }
        }

        result.add(level);
        leftToRight = !leftToRight;
    }

    return result;
}
```


## Maximum Depth of Binary Tree {#lc-max-depth}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 104
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/maximum-depth-of-binary-tree/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Return the maximum depth (number of nodes along the longest path from root down to a leaf).

### Examples

**Example 1: Balanced Tree**
```
Input:
    3
   / \
  9  20
    /  \
   15   7
Output: 3
Explanation: Longest path is 3 → 20 → 15 (or 3 → 20 → 7), which has 3 nodes
```

**Example 2: Single Node**
```
Input: 1
Output: 1
```

**Example 3: Skewed Tree**
```
Input:
  1
   \
    2
     \
      3
Output: 3
Explanation: Linear right spine, depth equals number of nodes
```

### Constraints

- The number of nodes in the tree is in the range `[0, 10^4]`
- `-100 <= Node.val <= 100`

### Solution - DFS Recursion

#### Walkthrough

The maximum depth is the longest path from root to any leaf node. Use a recursive DFS approach:

1. **Base case**: If current node is null, return 0 (no depth).
2. **Recursive case**: The depth of the current node is 1 (for the node itself) plus the maximum of the depths of its left and right subtrees.
3. This naturally explores all paths and returns the longest one.

The elegance of this approach is that max(left_depth, right_depth) inherently finds the longest path at each level as it bubbles up the recursion stack.

#### Analysis

- **Time Complexity**: O(n) — each node is visited exactly once
- **Space Complexity**: O(h) — recursion stack depth, where h is tree height (O(n) in worst case for skewed tree)

#### Implementation Steps

1. Handle null node as base case returning 0
2. Recursively calculate depth of left subtree
3. Recursively calculate depth of right subtree
4. Return 1 + max(left_depth, right_depth)

#### Code - Java

```
public int maxDepth(TreeNode root) {
    if (root == null) {
        return 0;
    }
    return 1 + Math.max(maxDepth(root.left), maxDepth(root.right));
}
```


## Construct Binary Tree from Preorder and Inorder Traversal {#lc-build-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 105
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/construct-binary-tree-from-preorder-and-inorder-traversal/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Divide and Conquer](glossary-of-algorithm.html#glossary-dnc), [Recursion](glossary-of-algorithm.html#glossary-recursion), [Hash Table](glossary-of-algorithm.html#glossary-hash-table), Tree

### Description

Build a binary tree from preorder and inorder traversals.

### Examples

**Example 1: Balanced Tree**
```
Input: preorder = [3,9,20,15,7], inorder = [9,3,15,20,7]
Output:
    3
   / \
  9  20
    /  \
   15   7
Explanation:
- Preorder: root first (3), then left subtree (9,20,15,7), ... wait that's wrong
- Preorder: 3 (root), 9 (left subtree root), 20 (right subtree root), 15, 7
- Inorder: 9 (left), 3 (root), 15,20,7 (right)
```

**Example 2: Single Node**
```
Input: preorder = [1], inorder = [1]
Output: 1
```

**Example 3: Left-Skewed Tree**
```
Input: preorder = [1,2,3], inorder = [3,2,1]
Output:
  1
 /
2
/
3
```

### Constraints

- `1 <= preorder.length <= 3000`
- `inorder.length == preorder.length`
- `-3000 <= preorder[i], inorder[i] <= 3000`
- `preorder` and `inorder` consist of unique values
- Each value of `inorder` also appears in `preorder`
- `preorder` is guaranteed to be a valid preorder traversal
- `inorder` is guaranteed to be a valid inorder traversal

### Solution - Hash Map with Divide and Conquer

#### Walkthrough

The key insight is that **the first element of preorder is always the root**. Once we find the root in inorder, we can divide the remaining elements into left and right subtrees:

1. **Preorder property**: First element is always the root; remaining elements are [left subtree elements, right subtree elements]
2. **Inorder property**: Elements are [left subtree elements, root, right subtree elements]
3. **Algorithm**:
   - Take first element of preorder as root
   - Find its position in inorder
   - Elements to the left of root in inorder form the left subtree
   - Elements to the right of root in inorder form the right subtree
   - Recursively build left and right subtrees using the corresponding preorder and inorder subarrays
4. **Optimization**: Use a HashMap to store inorder indices for O(1) root position lookup

#### Analysis

- **Time Complexity**: O(n) — each node is visited once, and HashMap lookup is O(1)
- **Space Complexity**: O(n) — HashMap stores all n values; recursion stack is O(h) where h is tree height

#### Implementation Steps

1. Build a HashMap: `value → index` from the inorder array for fast lookup
2. Define a recursive helper function that takes preorder bounds and inorder bounds
3. In each recursion:
   - Extract root value from preorder[preL]
   - Find root position in inorder using HashMap
   - Calculate left subtree size = (root position in inorder) - inL
   - Recursively build left subtree: preorder[preL+1 to preL+leftSize], inorder[inL to rootPos-1]
   - Recursively build right subtree: preorder[preL+leftSize+1 to preR], inorder[rootPos+1 to inR]
4. Base case: when preL > preR, return null

#### Code - Java

```java
public TreeNode buildTree(int[] preorder, int[] inorder) {
    Map<Integer, Integer> indexMap = new HashMap<>();
    for (int i = 0; i < inorder.length; i++) {
        indexMap.put(inorder[i], i);
    }
    return build(preorder, 0, preorder.length - 1, 0, inorder.length - 1, indexMap);
}

private TreeNode build(int[] preorder, int preL, int preR, int inL, int inR, Map<Integer, Integer> indexMap) {
    if (preL > preR) {
        return null;
    }
    int rootVal = preorder[preL];
    TreeNode root = new TreeNode(rootVal);
    int inorderRootIndex = indexMap.get(rootVal);
    int leftSize = inorderRootIndex - inL;
    root.left = build(preorder, preL + 1, preL + leftSize, inL, inorderRootIndex - 1, indexMap);
    root.right = build(preorder, preL + leftSize + 1, preR, inorderRootIndex + 1, inR, indexMap);
    return root;
}
```

## Binary Tree Level Order Traversal II {#lc-level-order-bottom}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 107
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-level-order-traversal-ii/
- **Tags**:
- **Techniques**: [Breadth First Search](#intro-tree-bfs), Tree

### Description

Return the level order traversal of a binary tree from the bottom level up to the root.

### Examples

```
Input: [3,9,20,null,null,15,7]
Output: [[15,7],[9,20],[3]]
```

### Constraints

- `0 <= nodes <= 2000`

### Solution - Reverse BFS

#### Walkthrough

Perform BFS collecting nodes level by level. Instead of appending to the end, insert each level at the front of a linked list (or append and reverse at the end).

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Implementation Steps

1. Run BFS using a queue.
2. For each level size, dequeue nodes, record values, enqueue children.
3. Insert level list at the start of the result.

#### Code - Java
```java
public List<List<Integer>> levelOrderBottom(TreeNode root) {
    LinkedList<List<Integer>> result = new LinkedList<>();
    if (root == null) {
        return result;
    }

    Queue<TreeNode> queue = new LinkedList<>();
    queue.offer(root);

    while (!queue.isEmpty()) {
        int size = queue.size();
        List<Integer> level = new ArrayList<>(size);

        for (int i = 0; i < size; i++) {
            TreeNode node = queue.poll();
            level.add(node.val);
            if (node.left != null) {
                queue.offer(node.left);
            }
            if (node.right != null) {
                queue.offer(node.right);
            }
        }

        result.addFirst(level);
    }

    return result;
}
```


## Convert Sorted Array to BST {#lc-sorted-array-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 108
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/convert-sorted-array-to-binary-search-tree/
- **Tags**:
- **Techniques**: [Divide and Conquer](glossary-of-algorithm.html#glossary-dnc), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Convert a sorted array to a height-balanced BST.

### Examples

**Example 1: Array with Odd Length**
```
Input: nums = [-10,-3,0,5,9]
Output:
    0
   / \
 -3   9
 /   /
-10 5
Explanation: Middle element (0) becomes root; recurse on left and right halves
```

**Example 2: Array with Even Length**
```
Input: nums = [1,3]
Output:
  3
 /
1
(or vice versa, both are valid height-balanced BSTs)
```

**Example 3: Single Element**
```
Input: nums = [1]
Output: 1
```

### Constraints

- `1 <= nums.length <= 10^4`
- `-10^4 <= nums[i] <= 10^4`
- `nums` is sorted in strictly increasing order

### Solution - Divide and Conquer

#### Walkthrough

To build a height-balanced BST from a sorted array, use the middle element as the root. This ensures the tree remains balanced since it divides the remaining elements equally between left and right subtrees:

1. **Base case**: If left > right, return null (no elements).
2. **Find middle**: Calculate mid = left + (right - left) / 2 to avoid integer overflow.
3. **Create root**: Use nums[mid] as the current node.
4. **Recurse left**: Recursively build left subtree from nums[left...mid-1].
5. **Recurse right**: Recursively build right subtree from nums[mid+1...right].
6. The middle element selection ensures all left subtree values < root < all right subtree values (BST property).

#### Analysis

- **Time Complexity**: O(n) — each element is visited once
- **Space Complexity**: O(log n) — recursion stack depth for balanced tree (O(n) in worst case)

#### Implementation Steps

1. Define a recursive helper function with left and right indices
2. Handle base case when left > right
3. Calculate middle index
4. Create node with middle value
5. Recursively build left subtree (left to mid-1)
6. Recursively build right subtree (mid+1 to right)
7. Return the constructed node

#### Code - Java

```
public TreeNode sortedArrayToBST(int[] nums) {
    return build(nums, 0, nums.length - 1);
}

private TreeNode build(int[] nums, int left, int right) {
    if (left > right) {
        return null;
    }
    int mid = left + (right - left) / 2;
    TreeNode node = new TreeNode(nums[mid]);
    node.left = build(nums, left, mid - 1);
    node.right = build(nums, mid + 1, right);
    return node;
}
```


## Balanced Binary Tree {#lc-balanced-binary-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 110
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/lc-balanced-binary-tree/
- **Tags**: Grind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree, Binary Tree

### Description

Given a binary tree, determine if it is **height-balanced**.

For this problem, a height-balanced binary tree is defined as:

> A binary tree in which the depth of the two subtrees of **every** node never differ by more than 1.

### Examples

**Example 1:**
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Balanced tree example', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[
level 1/.style={sibling distance=10em},
level 2/.style={sibling distance=5em},
level 3/.style={sibling distance=5em},
level distance=2.5em,
]
\node {3}
    child { node {9}}
    child { node {20}
        child { node {15} }
        child { node {7} }
    }
;
\end{tikzpicture}
```
```
Input: root = [3,9,20,null,null,15,7]
Output: true
Explanation: Height of left subtree = 1, height of right subtree = 2, difference = 1 (balanced)
```

**Example 2:**
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Unbalanced tree example', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[
    level 1/.style={sibling distance=10em},
    level 2/.style={sibling distance=5em},
    level 3/.style={sibling distance=5em},
    level distance=2.5em,
    ]
\node {1}
    child { node {2}
        child {  node {3}
            child {  node {4}}
            child {  node {4}}
        }
        child {  node {3}}
    }
    child { node {2}}
;
\end{tikzpicture}
```
```
Input: root = [1,2,2,3,3,null,null,4,4]
Output: false
Explanation: Left subtree has much greater depth than right subtree
```

**Example 3:**
```
Input: root = []
Output: true
```

### Constraints

- The number of nodes in the tree is in the range `[0, 5000]`
- `-10^4 <= Node.val <= 10^4`

### Solution - Recursive DFS

#### Walkthrough

Use a recursive depth-first search to compute the height of each subtree.

For each node:
1. Recursively compute the height of the left subtree
2. Recursively compute the height of the right subtree
3. Check if the absolute difference between heights is greater than 1
4. If unbalanced at any node, propagate `-1` upward to indicate imbalance
5. Otherwise, return the maximum height + 1

The key insight is to combine height calculation with balance checking in a single pass.

#### Analysis

- **Time Complexity**: O(n) - Each node is visited exactly once
- **Space Complexity**: O(h) - Recursion call stack depth equals tree height h, where h can be log(n) for balanced trees or n for skewed trees

#### Implementation Steps


1. If root is null, return 0 (base case)
2. Recursively get height of left subtree: `lHeight = getHeight(root.left)`
3. Recursively get height of right subtree: `rHeight = getHeight(root.right)`
4. If either subtree returned -1 (unbalanced), propagate -1 upward
5. If `|lHeight - rHeight| > 1`, return -1 (unbalanced)
6. Otherwise, return `max(lHeight, rHeight) + 1`

#### Code - Java

```java
public boolean isBalanced(TreeNode root) {
    return getHeight(root) != -1;
}

private int getHeight(TreeNode root) {
    if (root == null) {
        return 0;
    }

    int lHeight = getHeight(root.left);
    int rHeight = getHeight(root.right);

    // If left or right subtree is unbalanced, propagate -1
    if (lHeight == -1 || rHeight == -1) {
        return -1;
    }

    // If current node is unbalanced, return -1
    if (Math.abs(lHeight - rHeight) > 1) {
        return -1;
    }

    // Return height of current subtree
    return Math.max(lHeight, rHeight) + 1;
}
```

#### Code - Golang

```go
func isBalanced(root *TreeNode) bool {
    if root == nil {
        return true
    }

    return dfs(root) != -1
}

func dfs(node *TreeNode) int {
    if node == nil {
        return 0
    }

    lHeight := dfs(node.Left)
    rHeight := dfs(node.Right)

    // If left or right subtree is unbalanced, return -1
    if lHeight == -1 || rHeight == -1 {
        return -1
    }

    // If current node is unbalanced, return -1
    if math.Abs(float64(lHeight - rHeight)) > 1 {
        return -1
    }

    // return height of current subtree
    return max(lHeight, rHeight) + 1
}
```

## Binary Tree Level Order Traversal {#lc-level-order}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 102
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-level-order-traversal/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Breadth First Search](#intro-tree-bfs), Tree

### Description
Given the root of a binary tree, return the level order traversal of its nodes' values (i.e., from left to right, level by level).

### Examples

**Example 1:**
Input: root = [3,9,20,null,null,15,7]
Output: [[3],[9,20],[15,7]]

**Example 2:**
Input: root = [1]
Output: [[1]]

**Example 3:**
Input: root = []
Output: []

### Constraints
- The number of nodes in the tree is in the range `[0, 2000]`
- `-1000 <= Node.val <= 1000`

### Solution - BFS with Two Queues
#### Walkthrough
Use level-order traversal while keeping two queues: one for the current level, one for the next.

1. Seed `currentLevelQueue` with root.
2. Pop nodes from the current queue, append their values to the `level` list, and push their children into `nextLevelQueue`.
3. When the current queue empties, swap queues, reset `level`, and continue; each swap marks a new level in `levels`.

This separates nodes by depth without tracking counts.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n) for queues and result

#### Implementation Steps

1. Initialize two queues and an empty result list.
2. While current queue not empty, pop nodes, append values, and enqueue children to the next queue.
3. When current queue empties, swap queues and start a new level list.

#### Code - Java
```java
public List<List<Integer>> levelOrder(TreeNode root) {
    List<List<Integer>> levels = new ArrayList<>();
    if (root == null) {
        return levels;
    }

    // Two queues for current level and next level
    Queue<TreeNode> currLvlQ = new LinkedList<>();
    Queue<TreeNode> nextLvlQ = new LinkedList<>();
    // One list for storing current level values
    List<Integer> level = new ArrayList<>();

    // Enqueue root
    currLvlQ.offer(root);

    while (!currLvlQ.isEmpty()) {
        // Dequeue from current level
        TreeNode node = currLvlQ.poll();

        // Add node value to current level list
        level.add(node.val);

        // Enqueue left/right children to next level queue
        if (node.left != null) {
            nextLvlQ.offer(node.left);
        }
        if (node.right != null) {
            nextLvlQ.offer(node.right);
        }

        // Proceed to next level when current queue is empty
        if (currLvlQ.isEmpty()) {
            // Replace currLvlQ with nextLvlQ
            currLvlQ = nextLvlQ;
            nextLvlQ = new LinkedList<>();

            // Add current level to result 2D list
            levels.add(level);

            // Initialize new level list
            level = new ArrayList<>();
        }
    }

    return levels;
}
```

#### Code - Golang
```go
func levelOrder(root *TreeNode) [][]int {
    levels := make([][]int, 0)

    if root == nil {
        return levels
    }

    // Two queues for current level and next level
    currLvlQ := make([]*TreeNode, 0)
    nextLvlQ := make([]*TreeNode, 0)
    // One list for storing current level values
    level := make([]int, 0)

    // Enqueue root
    currLvlQ = append(currLvlQ, root)

    for len(currLvlQ) != 0 {
        // Dequeue from current level
        node := currLvlQ[0]
        currLvlQ = currLvlQ[1:]

        // Add node value to current level list
        level = append(level, node.Val)

        // Enqueue left/right children to next level queue
        if node.Left != nil {
            nextLvlQ = append(nextLvlQ, node.Left)
        }
        if node.Right != nil {
            nextLvlQ = append(nextLvlQ, node.Right)
        }

        // Proceed to next level when current queue is empty
        if len(currLvlQ) == 0 {
            // Replace currLvlQ with nextLvlQ
            currLvlQ = nextLvlQ
            nextLvlQ = make([]*TreeNode, 0)

            // Add current level to result 2D list
            levels = append(levels, level)

            // Initialize new level list
            level = make([]int, 0)
        }
    }

    return levels
}
```

### Solution - DFS
#### Walkthrough
We could traverse the tree recursively with \textbf{PreOrder} strategy, and additionally pass current level
into recursive method.

#### Analysis
DFS has time complexity of O(n) as every node is visited and thus Auxiliary Space is also O(n).

#### Implementation Steps

 dfs \@ref(glossary-dfs)

#### Code - DFS
```java
public List<List<Integer>> levelOrder(TreeNode root) {
    List<List<Integer>> ans = new ArrayList<>();
    traverse(root, 0, ans);

    return ans;
}

private void traverse(TreeNode root, int level, List<List<Integer>> ans) {
    if(root == null){
        return;
    }

    if(ans.size()<=level){
        ans.add(new ArrayList<>());
    }

    ans.get(level).add(root.val);

    traverse(root.left, level + 1, ans);
    traverse(root.right, level + 1, ans);
}
```


## Minimum Depth of Binary Tree {#lc-min-depth}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 111
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/minimum-depth-of-binary-tree/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), DFS, Tree

### Description

Return the minimum depth from root to the nearest leaf.

### Solution - DFS

#### Walkthrough

Return 0 for a null node. If one child is null, the minimum depth must go through the other child. Otherwise, take `1 + min(leftDepth, rightDepth)`.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java
```
public int minDepth(TreeNode root) {
    if (root == null) {
        return 0;
    }
    if (root.left == null) {
        return 1 + minDepth(root.right);
    }
    if (root.right == null) {
        return 1 + minDepth(root.left);
    }
    return 1 + Math.min(minDepth(root.left), minDepth(root.right));
}
```


## Path Sum {#lc-path-sum}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 112
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/lc-path-sum/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), Tree

### Description

Determine if a binary tree has a root-to-leaf path whose node values sum to a target.

### Examples

```
Input: root = [5,4,8,11,null,13,4,7,2,null,null,null,1], target = 22
Output: true
```

### Constraints

- `0 <= nodes <= 5000`

### Solution - DFS with Running Sum

#### Walkthrough

Traverse the tree, accumulating the current sum along the path. When a leaf is reached, check if the sum equals the target. Short-circuit as soon as a valid path is found.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java
```
public boolean hasPathSum(TreeNode root, int targetSum) {
    if (root == null) {
        return false;
    }
    if (root.left == null && root.right == null) {
        return targetSum == root.val;
    }
    int remaining = targetSum - root.val;
    return hasPathSum(root.left, remaining) || hasPathSum(root.right, remaining);
}
```


## Path Sum II {#lc-path-sum-ii}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 113
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lc-path-sum-ii/
- **Tags**:
- **Techniques**: [Backtracking: Choose-Explore-Unchoose](#backtracking-choose-explore-unchoose-pattern), Tree, [Backtracking](glossary-of-algorithm.html#glossary-backtracking)

### Description

Find all root-to-leaf paths where the sum of node values equals a target.

### Solution - Backtracking

#### Walkthrough

This problem uses the **2-dimensional traversal mental model** of backtracking (see [Pattern 1: Choose-Explore-Unchoose](#backtracking-choose-explore-unchoose-pattern)):

- **Outer dimension** (recursion parameter): `node` / depth in the tree - controls how deep we've traversed
- **Inner dimension** (implicit traversal): At each node, we have up to 2 choices (left child and right child)

The backtracking explores the tree as a conceptual decision space where each level represents a depth in the tree, and at each level we "choose" to include the current node's value in the path.

**Key mechanism**: Traverse the tree, maintaining the current path list and remaining sum. When a leaf reaches zero remaining sum, append a copy of the path to the answer. Backtrack after exploring each branch to try other paths.

#### Analysis

- **Time Complexity**: O(n * h)
- **Space Complexity**: O(h)

#### Code - Java
```
public List<List<Integer>> pathSum(TreeNode root, int targetSum) {
    List<List<Integer>> result = new ArrayList<>();
    backtrack(root, targetSum, new ArrayList<>(), result);
    return result;
}

private void backtrack(TreeNode node, int remain, List<Integer> path, List<List<Integer>> result) {
    if (node == null) {
        return;
    }

    path.add(node.val);
    remain -= node.val;

    if (node.left == null && node.right == null && remain == 0) {
        result.add(new ArrayList<>(path));
    } else {
        backtrack(node.left, remain, path, result);
        backtrack(node.right, remain, path, result);
    }

    path.remove(path.size() - 1);
}
```


## Populating Next Right Pointers in Each Node {#lc-connect-next}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 116
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/populating-next-right-pointers-in-each-node/
- **Tags**:
- **Techniques**: [Breadth First Search](#intro-tree-bfs), Tree

### Description

Connect each node's `next` pointer to its right neighbor on the same level in a perfect binary tree.

### Solution - Level Traversal

#### Walkthrough

Use the `next` pointers to traverse each level without a queue. For each node, connect `left -> right`, and if a `next` neighbor exists, connect `right -> neighbor.left`. Then advance to the leftmost node of the next level.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(1)

#### Code - Java
```
public Node connect(Node root) {
    if (root == null) {
        return null;
    }
    Node leftmost = root;
    while (leftmost.left != null) {
        Node head = leftmost;
        while (head != null) {
            head.left.next = head.right;
            if (head.next != null) {
                head.right.next = head.next.left;
            }
            head = head.next;
        }
        leftmost = leftmost.left;
    }
    return root;
}
```


## Binary Tree Preorder Traversal {#lc-preorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 144
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/binary-tree-preorder-traversal/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Pre-Order Traversal](#intro-tree-preorder), [Stack](glossary-of-algorithm.html#glossary-stack), Tree

### Description

Return the preorder traversal (root-left-right) of a binary tree.

### Solution - Iterative Stack

#### Walkthrough

Push root to a stack. Pop, record value, then push right and left children (right first so left is processed next). Repeat until stack empty.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n)

#### Code - Java
```java
public List<Integer> preorderTraversal(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    if (root == null) {
        return result;
    }

    Deque<TreeNode> stack = new ArrayDeque<>();
    stack.push(root);

    while (!stack.isEmpty()) {
        TreeNode node = stack.pop();
        result.add(node.val);
        if (node.right != null) {
            stack.push(node.right);
        }
        if (node.left != null) {
            stack.push(node.left);
        }
    }
    return result;
}
```

### Solution - Recursive

#### Walkthrough

Recursively visit the current node first, then traverse left subtree, then right subtree. This mirrors the preorder definition and appends values in root-left-right order.

The recursive approach directly reflects the definition of pre-order traversal:
1. **Base case**: If node is null, return immediately
2. **Process current**: Add current node's value to result
3. **Recurse left**: Visit all nodes in the left subtree
4. **Recurse right**: Visit all nodes in the right subtree

This is the natural traversal order for creating tree copies or serializing trees, as you process the parent before the children.

#### Analysis

- **Time Complexity**: O(n) - Each node is visited exactly once
- **Space Complexity**: O(h) - Recursion call stack depth where h is tree height (O(n) worst case for skewed tree)

#### Code - Java

```java
public List<Integer> preorderTraversalRecursive(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    dfs(root, result);
    return result;
}

private void dfs(TreeNode node, List<Integer> list) {
    if (node == null) {
        return;
    }
    list.add(node.val);
    dfs(node.left, list);
    dfs(node.right, list);
}
```


## Binary Tree Postorder Traversal {#lc-postorder-traversal}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 145
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/binary-tree-postorder-traversal/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Post-Order Traversal](#intro-tree-postorder), [Stack](glossary-of-algorithm.html#glossary-stack), Tree

### Description

Return the postorder traversal (left-right-root) of a binary tree.

### Solution - Iterative Stack

#### Walkthrough

Use a reversed-preorder trick to achieve postorder without two stacks:

1. Push root.
2. Pop node, prepend its value (`addFirst`) to result.
3. Push left then right children (so right is processed before left).
4. Prepending reverses the root-right-left traversal into left-right-root.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(n) for stack + output list

```java
public List<Integer> postorderTraversal(TreeNode root) {
    LinkedList<Integer> result = new LinkedList<>();
    if (root == null) {
        return result;
    }

    Deque<TreeNode> stack = new ArrayDeque<>();
    stack.push(root);
    while (!stack.isEmpty()) {
        TreeNode node = stack.pop();
        result.addFirst(node.val);
        if (node.left != null) {
            stack.push(node.left);
        }
        if (node.right != null) {
            stack.push(node.right);
        }
    }
    return result;
}
```

### Solution - Recursive

#### Walkthrough

Recursively traverse left subtree, then right subtree, then visit the current node. This mirrors the postorder definition and appends values in left-right-root order.

The recursive approach directly reflects the definition of post-order traversal:
1. **Base case**: If node is null, return immediately
2. **Recurse left**: Visit all nodes in the left subtree
3. **Recurse right**: Visit all nodes in the right subtree
4. **Process current**: Add current node's value to result

This is the natural traversal order for deleting trees or performing bottom-up calculations, as children are processed before their parent.

#### Analysis

- **Time Complexity**: O(n) - Each node is visited exactly once
- **Space Complexity**: O(h) - Recursion call stack depth where h is tree height (O(n) worst case for skewed tree)

#### Code - Java

```java
public List<Integer> postorderTraversalRecursive(TreeNode root) {
    List<Integer> result = new ArrayList<>();
    dfs(root, result);
    return result;
}

private void dfs(TreeNode node, List<Integer> list) {
    if (node == null) {
        return;
    }
    dfs(node.left, list);
    dfs(node.right, list);
    list.add(node.val);
}
```


## Implement Trie (Prefix Tree) {#lc-implement-trie}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 208
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/implement-trie-prefix-tree/
- **Tags**: Grind 75, NeetCode 150
- **Techniques**: [Trie](glossary-of-algorithm.html#glossary-trie), Tree, [Hash Table](glossary-of-algorithm.html#glossary-hash-table)

### Description

A **trie** (pronounced as "try") or **prefix tree** is a tree data structure used to efficiently store and retrieve keys in a dataset of strings. There are various applications of this data structure, such as autocomplete and spellchecker.

Implement the Trie class:

- `Trie()` Initializes the trie object.
- `void insert(String word)` Inserts the string `word` into the trie.
- `boolean search(String word)` Returns `true` if the string `word` is in the trie (i.e., was inserted before), and `false` otherwise.
- `boolean startsWith(String prefix)` Returns `true` if there is a previously inserted string `word` that has the prefix `prefix`, and `false` otherwise.

### Examples

**Example 1:**
```
Input:
["Trie", "insert", "search", "search", "startsWith", "insert", "search"]
[[], ["apple"], ["apple"], ["app"], ["app"], ["app"], ["app"]]

Output:
[null, null, true, false, true, null, true]

Explanation:
Trie trie = new Trie();
trie.insert("apple");
trie.search("apple");   // return True
trie.search("app");     // return False
trie.startsWith("app"); // return True
trie.insert("app");
trie.search("app");     // return True
```

### Constraints

- `1 <= word.length, prefix.length <= 2000`
- `word` and `prefix` consist only of lowercase English letters.
- At most $3 \times 10^4$ calls in total will be made to `insert`, `search`, and `startsWith`.

### Solution - TrieNode with HashMap

#### Walkthrough

A Trie is a tree-like data structure where each node represents a character. The key insight is that words sharing common prefixes share the same path from the root.

**Data Structure Design:**

1. **TrieNode structure**: Each node contains:
   - A `Map<Character, TrieNode>` to store child nodes (one for each possible next character)
   - A `boolean isEndOfWord` flag to mark if this node completes a valid word

2. **Root node**: Empty node serving as the entry point for all words

**Algorithm for each operation:**

**Insert(word):**
1. Start at root node
2. For each character in word:
   - If character doesn't exist as a child, create new TrieNode
   - Move to child node
3. Mark final node as end of word

**Search(word):**
1. Start at root node
2. For each character in word:
   - If character doesn't exist as a child, return false
   - Move to child node
3. Return true only if final node is marked as end of word

**StartsWith(prefix):**
1. Start at root node
2. For each character in prefix:
   - If character doesn't exist as a child, return false
   - Move to child node
3. Return true (we've found the complete prefix path)

**Why HashMap for children?**
- Supports only lowercase letters (26 possible children)
- HashMap provides O(1) lookup and insertion
- Alternative: array of size 26 (faster but wastes space for sparse tries)

#### Analysis

Let n = length of word/prefix being processed, m = total number of characters inserted

- **Time Complexity**:
  - `insert(word)`: O(n) - traverse each character once
  - `search(word)`: O(n) - traverse each character once
  - `startsWith(prefix)`: O(n) - traverse each character once

- **Space Complexity**: O(m) - In worst case, no words share common prefixes, requiring one node per character inserted

#### Implementation Steps

1. **Define TrieNode class**:
   - HashMap to store children nodes
   - Boolean flag for end of word marker

2. **Initialize Trie**:
   - Create root node (empty node)

3. **Implement insert**:
   - Start from root
   - For each character, create node if not exists
   - Navigate to child node
   - Mark final node as word ending

4. **Implement search**:
   - Traverse character by character
   - Return false if path breaks
   - Check end-of-word flag at final node

5. **Implement startsWith**:
   - Same as search but skip end-of-word check
   - Return true if complete prefix path exists

#### Code - Java

```java
class Trie {
    private class TrieNode {
        Map<Character, TrieNode> children;
        boolean isEndOfWord;

        TrieNode() {
            children = new HashMap<>();
            isEndOfWord = false;
        }
    }

    private TrieNode root;

    public Trie() {
        root = new TrieNode();
    }

    public void insert(String word) {
        TrieNode current = root;
        for (char ch : word.toCharArray()) {
            current.children.putIfAbsent(ch, new TrieNode());
            current = current.children.get(ch);
        }
        current.isEndOfWord = true;
    }

    public boolean search(String word) {
        TrieNode current = root;
        for (char ch : word.toCharArray()) {
            current = current.children.get(ch);
            if (current == null) {
                return false;
            }
        }
        return current.isEndOfWord;
    }

    public boolean startsWith(String prefix) {
        TrieNode current = root;
        for (char ch : prefix.toCharArray()) {
            current = current.children.get(ch);
            if (current == null) {
                return false;
            }
        }
        return true;
    }
}
```

### Solution - Array-Based TrieNode

#### Walkthrough

Instead of using a HashMap, we can use a **fixed-size array of 26 elements** to represent children nodes, where each index corresponds to a letter (0 for 'a', 1 for 'b', ..., 25 for 'z').

**Key Differences from HashMap approach:**

1. **Space Optimization**: Only allocate array when first child is added (lazy initialization)
2. **Direct Access**: Character mapping via `ch - 'a'` gives O(1) array index
3. **Better Cache Locality**: Array elements stored contiguously in memory

**Algorithm remains the same**, but child lookup becomes:
- HashMap: `children.get(ch)` → hash computation + bucket lookup
- Array: `letters[ch - 'a']` → direct array access

**Trade-offs:**
- **Pros**: Faster access (no hashing), better for small alphabets
- **Cons**: Wastes space if words are sparse (many empty array slots)

#### Analysis
Let n = length of word/prefix being processed, m = total number of characters inserted

- **Time Complexity**: Same as HashMap version - O(n) for all operations
- **Space Complexity**: O(m × 26) worst case, but often better in practice due to fewer object allocations

#### Code - Java (Array-Based, Recursive)

```java
class Trie {
    private Trie[] letters = new Trie[26];
    private boolean isEndOfWord = false;

    public Trie() {
    }

    public void insert(String word) {
        insertRec(word, 0, this);
    }

    private void insertRec(String word, int i, Trie node) {
        if (i >= 0 && i < word.length()) {
            int index = word.charAt(i) - 'a';

            // Insert a node at letter index if not exists
            if (node.letters[index] == null) {
                node.letters[index] = new Trie();
            }

            insertRec(word, i + 1, node.letters[index]);
        }

        if (i == word.length()) {
            node.isEndOfWord = true;
        }
    }

    public boolean search(String word) {
        Trie result = findRec(word, 0, this);
        return result != null && result.isEndOfWord;
    }

    public boolean startsWith(String prefix) {
        return findRec(prefix, 0, this) != null;
    }

    private Trie findRec(String word, int i, Trie node) {
        if (i >= 0 && i < word.length()) {
            int index = word.charAt(i) - 'a';

            // Fail to find letter at letter index
            if (node.letters[index] == null) {
                return null;
            }

            return findRec(word, i + 1, node.letters[index]);
        } else {
            // Leaf node
            return node;
        }
    }
}
```

**Implementation Notes:**

- **Recursive Design**: Uses helper methods `insertRec` and `findRec` for cleaner separation
- **Index Calculation**: `word.charAt(i) - 'a'` maps 'a' to 0, 'b' to 1, etc.
- **Base Case**: When `i == word.length()`, we've processed all characters
- **Null Check**: `letters[index] == null` indicates character path doesn't exist


## Kth Smallest Element in a BST {#lc-kth-smallest-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 230
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/kth-smallest-element-in-a-bst/
- **Tags**: Top 100 Liked
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Binary Search Tree](glossary-of-algorithm.html#glossary-bst), [In-Order Traversal](#intro-tree-inorder)

### Description

Given the root of a binary search tree, and an integer k, return the kth smallest value (1-indexed) of all the values of the nodes in the tree.

### Examples

**Example 1:**
```
Input: root = [3,1,4,null,2], k = 1
Tree:
    3
   / \
  1   4
   \
    2
Output: 1
Explanation: The in-order traversal is [1, 2, 3, 4], and the 1st smallest is 1
```

**Example 2:**
```
Input: root = [5,3,6,2,4,null,null,1], k = 3
Tree:
      5
     / \
    3   6
   / \
  2   4
 /
1
Output: 3
Explanation: The in-order traversal is [1, 2, 3, 4, 5, 6], and the 3rd smallest is 3
```

### Constraints

- The number of nodes in the tree is n
- $1 \le k \le n \le 10^4$
- $0 \le \text{Node.val} \le 10^4$

**Follow up:** If the BST is modified often (i.e., we can do insert and delete operations) and you need to find the kth smallest frequently, how would you optimize?

### Solution - In-Order Traversal (Recursive)

#### Walkthrough

The key insight is that **in-order traversal of a BST visits nodes in sorted ascending order** (left → root → right). Therefore, the kth smallest element is simply the kth node visited during in-order traversal.

**Algorithm:**

1. **Perform in-order DFS**: Traverse left subtree first, then process current node, then traverse right subtree
2. **Count visited nodes**: Maintain a counter that increments each time we visit a node
3. **Early termination**: When counter equals k, we've found the kth smallest element - store it and stop further traversal
4. **Use class-level state**: Store `count` and `result` as instance variables to track state across recursive calls

**Why In-Order Traversal Works:**

For a BST with the property that all left subtree values < root < all right subtree values, in-order traversal guarantees we visit nodes in ascending order:

```
Tree:    5
        / \
       3   6
      / \
     2   4
    /
   1

In-order visits: 1 → 2 → 3 → 4 → 5 → 6 (sorted!)
For k=3: Stop at the 3rd visit, which is node 3
```

**Optimization: Early Termination**

We don't need to visit all n nodes - we can stop as soon as we find the kth element. Check `result != -1` before recursing to avoid unnecessary traversals.

#### Analysis

- **Time Complexity**: O(H + k) where H is tree height
  - We traverse down to the leftmost node: O(H)
  - Then visit k nodes in in-order: O(k)
  - Total: O(H + k), which is O(k) for balanced trees since H = O(log n)
  - Best case: O(k) when tree is balanced
  - Worst case: O(n) when k = n or tree is skewed

- **Space Complexity**: O(H) for recursion stack
  - O(log n) for balanced tree
  - O(n) for skewed tree

#### Implementation Steps

1. Initialize class-level variables: `count = 0` and `result = -1`
2. Call in-order traversal helper starting from root
3. In the helper function:
   - Base case: if node is null or result already found, return
   - Recurse left subtree (visit smaller values first)
   - Increment count and check if count equals k
   - If count equals k, store node value in result and return
   - Recurse right subtree (visit larger values)
4. Return the result

#### Code - Java

```java
class Solution {
    private int count = 0;
    private int result = -1;

    public int kthSmallest(TreeNode root, int k) {
        inorder(root, k);
        return result;
    }

    private void inorder(TreeNode node, int k) {
        if (node == null || result != -1) {
            return;
        }

        // Left: visit all smaller values
        inorder(node.left, k);

        // Current: process this node
        count++;
        if (count == k) {
            result = node.val;
            return;
        }

        // Right: visit all larger values
        inorder(node.right, k);
    }
}
```

### Solution - Iterative with Stack

#### Walkthrough

An iterative approach using an explicit stack to simulate in-order traversal. This avoids recursion overhead and makes the traversal process more explicit.

**Algorithm:**

1. **Push all left children**: Starting from root, push nodes while moving left until we reach null
2. **Pop and process**: Pop a node from the stack (this is the next smallest value)
3. **Decrement k**: Each popped node represents one element in sorted order
4. **Check if done**: When k equals 0, we've found the answer
5. **Move to right subtree**: After processing a node, move to its right child and repeat

**Why This Works:**

The stack ensures we visit the leftmost (smallest) unvisited node next. After visiting a node, we explore its right subtree, which may contain values between this node and its parent.

```
Tree:    3
        / \
       1   4
        \
         2

Iteration steps:
1. Push 3, 1 (going left)
2. Pop 1 (1st smallest), k=1, continue
3. Push 2 (1.right)
4. Pop 2 (2nd smallest), k=0 → Return 2
```

#### Analysis

- **Time Complexity**: O(H + k)
  - Same as recursive: traverse to leftmost node (H), then visit k nodes

- **Space Complexity**: O(H) for the explicit stack
  - More predictable than recursion (no call stack overhead)

#### Code - Java

```java
class Solution {
    public int kthSmallest(TreeNode root, int k) {
        Deque<TreeNode> stack = new ArrayDeque<>();
        TreeNode current = root;

        while (current != null || !stack.isEmpty()) {
            // Push all left children
            while (current != null) {
                stack.push(current);
                current = current.left;
            }

            // Pop the next smallest node
            current = stack.pop();
            k--;

            // Found the kth smallest
            if (k == 0) {
                return current.val;
            }

            // Move to right subtree
            current = current.right;
        }

        return -1; // Should never reach here given constraints
    }
}
```

**Implementation Notes:**

- **Cleaner loop structure**: The outer while loop continues until we've explored all necessary nodes
- **No class-level state**: All state is local to the method
- **Better for production**: Iterative solutions are often easier to debug and reason about


## Lowest Common Ancestor of a Binary Search Tree {#lc-lca-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 235
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lowest-common-ancestor-of-a-binary-search-tree/
- **Tags**:
- **Techniques**: [Binary Search Tree](glossary-of-algorithm.html#glossary-bst), [Depth First Search](#intro-tree-dfs), Tree

### Description

Given a binary search tree (BST), find the lowest common ancestor (LCA) node of two given nodes in the BST.

The lowest common ancestor is defined as the lowest node in T that has both p and q as descendants (where we allow a node to be a descendant of itself).

### Examples

**Example 1:**
```
Input: root = [6,2,8,0,4,7,9,null,null,3,5], p = 2, q = 8
Output: 6
Explanation: The LCA of nodes 2 and 8 is 6.
```

**Example 2:**
```
Input: root = [6,2,8,0,4,7,9,null,null,3,5], p = 2, q = 4
Output: 2
Explanation: The LCA of nodes 2 and 4 is 2, since a node can be a descendant of itself according to the LCA definition.
```

**Example 3:**
```
Input: root = [2,1], p = 2, q = 1
Output: 2
```

### Constraints

- The number of nodes in the tree is in the range $[2, 10^5]$
- $-10^9 \le \text{Node.val} \le 10^9$
- All Node.val are unique
- $p \ne q$
- p and q will exist in the BST

### Solution - DFS with Ancestor Checking

#### Walkthrough

This is a general binary tree solution that works by recursively checking if nodes are ancestors:

1. **Base case**: If root is null, p, or q, return root
2. **Check if root is LCA**: Test if p and q are in opposite subtrees using `isAncestor()` helper
   - If `p` is in left subtree and `q` is in right subtree → root is LCA
   - If `p` is in right subtree and `q` is in left subtree → root is LCA
3. **Recursive search**: If root is not the LCA, recursively search both subtrees
4. **Return result**: Return the non-null result from either subtree

**Note**: This solution works for any binary tree but does not leverage BST properties.

#### Analysis

- **Time Complexity**: O(n²) — For each of n nodes, `isAncestor()` may traverse the entire subtree (O(n)), resulting in quadratic time
- **Space Complexity**: O(h) — Recursion stack depth where h is tree height

#### Code - Java

```java
public TreeNode lowestCommonAncestor(TreeNode root, TreeNode p, TreeNode q) {
    return dfs(root, p, q);
}

// Return true if root is an ancestor of node
private boolean isAncestor(TreeNode root, TreeNode node) {
    if (root == null) {
        return false;
    }

    if (root == node) {
        return true;
    }

    return isAncestor(root.left, node) || isAncestor(root.right, node);
}

private TreeNode dfs(TreeNode root, TreeNode p, TreeNode q) {
    if (root == null || root == p || root == q) {
        return root;
    }

    if (isAncestor(root.left, p) && isAncestor(root.right, q)) {
        return root;
    }

    if (isAncestor(root.right, p) && isAncestor(root.left, q)) {
        return root;
    }

    TreeNode lLCA = dfs(root.left, p, q);
    TreeNode rLCA = dfs(root.right, p, q);

    return lLCA == null ? rLCA : lLCA;
}
```

### Solution - BST Properties (Optimal)

#### Walkthrough

This solution leverages the BST property where all left subtree values are less than root and all right subtree values are greater than root:

1. **Compare both nodes with root value**:
   - If both `p.val` and `q.val` are **less than** `root.val` → LCA must be in the **left subtree**
   - If both `p.val` and `q.val` are **greater than** `root.val` → LCA must be in the **right subtree**
   - Otherwise → root is the **split point** (LCA)

2. **Split point cases**:
   - `p` and `q` are in different subtrees → root is their LCA
   - One of the nodes equals root → root is the LCA (a node can be its own ancestor)

3. **Recursive descent**: Follow the appropriate subtree until finding the split point

**Key Insight**: No need to search both subtrees - the BST ordering tells us exactly which direction to go.

#### Analysis

- **Time Complexity**: O(h) where h is tree height — Visit at most h nodes along the path from root to LCA (O(log n) for balanced BST, O(n) for skewed tree)
- **Space Complexity**: O(h) — Recursion stack depth

#### Implementation Steps

1. Compare both `p.val` and `q.val` with `root.val`
2. If both are smaller, recurse left
3. If both are larger, recurse right
4. Otherwise, return root as the LCA

#### Code - Java (Recursive)

```java
public TreeNode lowestCommonAncestor(TreeNode root, TreeNode p, TreeNode q) {
    // If both nodes are smaller, LCA is in left subtree
    if (p.val < root.val && q.val < root.val) {
        return lowestCommonAncestor(root.left, p, q);
    }

    // If both nodes are larger, LCA is in right subtree
    if (p.val > root.val && q.val > root.val) {
        return lowestCommonAncestor(root.right, p, q);
    }

    // Otherwise, root is the LCA (split point or one of the nodes)
    return root;
}
```

#### Code - Java (Iterative)

For even better space complexity, use an iterative approach:

```java
public TreeNode lowestCommonAncestor(TreeNode root, TreeNode p, TreeNode q) {
    while (root != null) {
        // Both nodes in left subtree
        if (p.val < root.val && q.val < root.val) {
            root = root.left;
        }
        // Both nodes in right subtree
        else if (p.val > root.val && q.val > root.val) {
            root = root.right;
        }
        // Found the split point
        else {
            return root;
        }
    }
    return null;
}
```

**Iterative Benefits**:
- **Space Complexity**: O(1) — No recursion stack
- **Time Complexity**: Still O(h)
- Cleaner and more efficient for BST-specific problem


## Two Sum IV - Input is a BST {#lc-two-sum-iv-bst}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 653
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/two-sum-iv-input-is-a-bst/
- **Tags**:
- **Techniques**: [Depth First Search](glossary-of-algorithm.html#glossary-dfs), [Hash Table](glossary-of-algorithm.html#glossary-hash-table), Tree, Binary Search Tree

### Description

Given the root of a Binary Search Tree and an integer `k`, return `true` if there exist **two distinct nodes** whose values sum to `k`. Otherwise, return `false`.

### Examples

```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='BST used in the example.', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[
level 1/.style={sibling distance=10em},
level 2/.style={sibling distance=5em},
level 3/.style={sibling distance=5em},
level distance=2.5em,
]
\node {5}
    child { node {3}
        child {  node {2} }
        child {  node {4} }
    }
    child { node {6}
        child {  node {7} }
    }
;
\end{tikzpicture}
```
```
Input: root = [5,3,6,2,4,null,7], k = 9
Output: true
Explanation: 2 + 7 = 9.
```

### Constraints

- `1 <= nodes <= 10^4`
- `-10^4 <= Node.val <= 10^4`
- BST property holds for the entire tree

### Solution - DFS + HashSet

#### Walkthrough

Traverse the tree with DFS while maintaining a set of seen node values. For each node:

1. Compute `needed = k - node.val`.
2. If `needed` is already in the set, return `true`.
3. Otherwise, insert `node.val` and continue exploring the left and right subtrees.

Because every value is checked exactly once, the traversal stays linear.

#### Analysis

- **Time Complexity**: O(n) - Visit each node exactly once
- **Space Complexity**: O(n) - HashSet stores up to n values + O(h) recursion stack where h is tree height

**Why Not Use BST Property?** While we could use in-order traversal to get sorted values and then apply two pointers, the DFS + HashSet approach is simpler and has the same O(n) time complexity.

**BST-Specific Optimization**: Could implement bidirectional BST iterator (one forward, one backward) for O(h) space instead of O(n), but added complexity rarely justifies the space savings.

**Trade-off**: We sacrifice O(n) space for code simplicity. The BST property doesn't provide significant advantage over treating it as a general binary tree for this problem.

#### Implementation Steps

1. Initialize an empty `HashSet<Integer>`.
2. Perform DFS:
   1. If the node is `null`, return `false`.
   2. If `set` contains `k - node.val`, return `true`.
   3. Add `node.val` to the set.
   4. Recurse into left or right subtrees; return `true` if either side finds a match.
3. Return `false` if no pair exists.

#### Code - Java

```java
public boolean findTarget(TreeNode root, int k) {
    return dfs(root, k, new HashSet<>());
}

private boolean dfs(TreeNode node, int target, Set<Integer> seen) {
    if (node == null) {
        return false;
    }

    int needed = target - node.val;
    if (seen.contains(needed)) {
        return true;
    }

    seen.add(node.val);
    return dfs(node.left, target, seen) || dfs(node.right, target, seen);
}
```

## Lowest Common Ancestor of a Binary Tree {#lc-lca-binary-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 236
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lowest-common-ancestor-of-a-binary-tree/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree, Binary Tree

### Description

Given a binary tree, find the lowest common ancestor (LCA) of two given nodes in the tree.

According to the definition of LCA on Wikipedia: "The lowest common ancestor is defined between two nodes `p` and `q` as the lowest node in `T` that has both `p` and `q` as descendants (where we allow **a node to be a descendant of itself**)."

### Examples

**Example 1:**
```
Input: root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 1
       3
      / \
     5   1
    / \ / \
   6  2 0  8
     / \
    7   4

Output: 3
Explanation: The LCA of nodes 5 and 1 is 3.
```

**Example 2:**
```
Input: root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 4
       3
      / \
     5   1
    / \ / \
   6  2 0  8
     / \
    7   4

Output: 5
Explanation: The LCA of nodes 5 and 4 is 5, since a node can be a descendant of itself according to the LCA definition.
```

**Example 3:**
```
Input: root = [1,2], p = 1, q = 2
  1
 /
2

Output: 1
```

### Constraints

- The number of nodes in the tree is in the range $[2, 10^5]$
- $-10^9 \le \text{Node.val} \le 10^9$
- All `Node.val` are **unique**
- $p \ne q$
- `p` and `q` will exist in the tree

### Solution 1 - DFS with Ancestor Checking

#### Walkthrough

This solution uses **explicit ancestor checking** to determine the LCA. The approach is straightforward but involves redundant tree traversals.

**Core Idea:**
For each node, explicitly check if `p` and `q` are in opposite subtrees using a helper function `isAncestor()`. The first node where this condition holds is the LCA.

**Algorithm Steps:**

1. **Base case**: If root is null, `p`, or `q`, return root immediately
2. **Check if root is LCA**: Use `isAncestor()` to test if `p` and `q` are in opposite subtrees:
   - If `p` is in left subtree AND `q` is in right subtree → root is LCA
   - If `p` is in right subtree AND `q` is in left subtree → root is LCA
3. **Recursive search**: If root is not the LCA, recursively search both subtrees
4. **Return result**: Return the non-null result from either subtree

**Helper Function `isAncestor(root, node)`:**
- Returns `true` if `node` exists in the subtree rooted at `root`
- Performs a complete DFS traversal of the subtree
- Checks both left and right children recursively

**Visual Example with root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 1:**

```
Tree structure:
       3
      / \
     5   1
    / \ / \
   6  2 0  8
     / \
    7   4

Step 1: dfs(3, 5, 1)
  - Check: isAncestor(3.left=5, p=5) → true (5 is in left subtree)
  - Check: isAncestor(3.right=1, q=1) → true (1 is in right subtree)
  - Both in opposite subtrees → Return 3 ✓

Result: 3 is the LCA
```

**Example with root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 4:**

```
Step 1: dfs(3, 5, 4)
  - Check: isAncestor(3.left=5, p=5) → true
  - Check: isAncestor(3.right=1, q=4) → false (4 not in right subtree)
  - Not in opposite subtrees, continue searching

Step 2: dfs(3.left=5, 5, 4)
  - Base case: root == p → Return 5 ✓

Result: 5 is the LCA (node can be ancestor of itself)
```

**Why This Works (But Inefficiently):**
- Guarantees correctness by explicitly checking ancestor relationships
- However, `isAncestor()` traverses entire subtrees multiple times
- For each node, we potentially traverse the tree multiple times

#### Analysis

- **Time Complexity**: O(n²)
  - For each of n nodes in the tree, we may call `isAncestor()`
  - Each `isAncestor()` call can traverse an entire subtree taking O(n) time
  - In worst case (skewed tree), we visit O(n) nodes, each doing O(n) work
  - Total: O(n × n) = O(n²)

- **Space Complexity**: O(h)
  - Recursion stack depth where h is tree height
  - `dfs()` recursion: O(h)
  - `isAncestor()` recursion nested within: O(h)
  - Combined maximum depth: O(h)
  - Best case (balanced tree): O(log n)
  - Worst case (skewed tree): O(n)

#### Implementation Steps

1. Define `isAncestor(root, node)` helper function:
   - Base case: if root is null, return `false`
   - If root equals node, return `true`
   - Recursively check if node exists in left OR right subtree

2. Define `dfs(root, p, q)` main recursive function:
   - Base case: if root is null, `p`, or `q`, return root
   - Check if `p` in left subtree AND `q` in right subtree → return root
   - Check if `p` in right subtree AND `q` in left subtree → return root
   - Recursively search left subtree for LCA
   - Recursively search right subtree for LCA
   - Return non-null result (left LCA if exists, otherwise right LCA)

3. Main function calls `dfs(root, p, q)` and returns result

#### Code - Java

```java
public TreeNode lowestCommonAncestor(TreeNode root, TreeNode p, TreeNode q) {
    return dfs(root, p, q);
}

// Return true if root is an ancestor of node
private boolean isAncestor(TreeNode root, TreeNode node) {
    if (root == null) {
        return false;
    }

    if (root == node) {
        return true;
    }

    return isAncestor(root.left, node) || isAncestor(root.right, node);
}

private TreeNode dfs(TreeNode root, TreeNode p, TreeNode q) {
    if (root == null || root == p || root == q) {
        return root;
    }

    if (isAncestor(root.left, p) && isAncestor(root.right, q)) {
        return root;
    }

    if (isAncestor(root.right, p) && isAncestor(root.left, q)) {
        return root;
    }

    TreeNode lLCA = dfs(root.left, p, q);
    TreeNode rLCA = dfs(root.right, p, q);

    return lLCA == null ? rLCA : lLCA;
}
```

### Solution 2 - Optimized Post-Order DFS

#### Walkthrough

This solution uses an **elegant recursive approach** that eliminates redundant tree traversals. The key insight is that we can determine the LCA in a single post-order traversal by leveraging the return values.

**Core Optimization:**
Instead of explicitly checking if nodes are ancestors, we use the recursive return values to implicitly communicate information about which nodes were found in each subtree.

**Key Insight:**
- When we find `p` or `q`, return it immediately
- As we bubble up the recursion:
  - If **both** left and right subtrees return non-null → current node is the LCA (split point)
  - If **only one** subtree returns non-null → propagate that result upward (LCA is higher up)
- This works because we're guaranteed `p` and `q` exist in the tree

**Algorithm Flow:**

1. **Base case**: If root is null OR root equals `p` OR root equals `q`, return root
   - Null case: no nodes found in this path
   - Match case: found one of the target nodes, return it

2. **Recursive exploration**: Search both subtrees
   - `left = lowestCommonAncestor(root.left, p, q)`
   - `right = lowestCommonAncestor(root.right, p, q)`

3. **Decision logic based on return values**:
   - **Both non-null** (`left != null` AND `right != null`) → found `p` in one subtree and `q` in the other → **root is LCA**
   - **Only left non-null** → both nodes are in left subtree, propagate left result upward
   - **Only right non-null** → both nodes are in right subtree, propagate right result upward
   - **Both null** → neither node found (shouldn't happen given constraints)

**Why This Works:**

The beauty of this solution lies in the **post-order traversal** nature:
- We process children before the parent
- Return values carry information: "Did I find p or q (or their LCA)?"
- When both subtrees return non-null, we know we've found the split point

**Visual Example with root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 1:**

```
Post-order traversal with return values:

                    3 → returns 3 (both left=5 and right=1 non-null)
                   / \
    returns 5 ← 5     1 → returns 1 (base case: root == q)
               / \   / \
              6   2 0   8
                 / \
                7   4

Execution trace:
1. Recurse left to node 5
2. Node 5 matches p → return 5 (base case)
3. Recurse right to node 1
4. Node 1 matches q → return 1 (base case)
5. Back at node 3: left=5 (non-null), right=1 (non-null)
6. Both non-null → node 3 is LCA → return 3
```

**Example with root = [3,5,1,6,2,0,8,null,null,7,4], p = 5, q = 4:**

```
                    3 → returns 5 (left=5, right=null)
                   / \
    returns 5 ← 5     1 → returns null
               / \   / \
              6   2 0   8
                 / \
    returns 4 ↑  7   4 ← base case returns 4

Execution trace:
1. Recurse left to node 5
2. Node 5 matches p → return 5 (but also need to check subtrees for q)
   - Wait, base case returns immediately! So return 5
3. Because of base case, we return 5 before checking its children
4. This works because the problem guarantees both nodes exist
5. At node 3: left=5 (non-null), right=null
6. Return left (which is 5) → 5 is the LCA
```

**Important Note on Base Case:**
The base case `if (root == p || root == q) return root` works because:
- We're guaranteed both `p` and `q` exist in the tree
- If we find `p` first, we know `q` is either in `p`'s subtree (making `p` the LCA) or not in this path
- The return value propagates upward, and the logic handles both cases correctly

#### Analysis

- **Time Complexity**: O(n)
  - Each node is visited exactly once in a single DFS traversal
  - No redundant tree traversals like in Solution 1
  - Best case: O(log n) if we find LCA near root in balanced tree
  - Worst case: O(n) if we must traverse entire tree
  - Average case: O(n) - must visit all nodes to ensure correctness

- **Space Complexity**: O(h)
  - Recursion call stack depth equals tree height h
  - Best case (balanced tree): O(log n)
  - Worst case (skewed tree): O(n)
  - No additional data structures needed

**Comparison with Solution 1:**
- Solution 1: O(n²) time due to repeated `isAncestor()` calls
- Solution 2: O(n) time with single traversal
- Both have O(h) space complexity
- Solution 2 is significantly more efficient for large trees

#### Implementation Steps

1. **Handle base cases**:
   - If root is null, return null (no nodes in empty tree)
   - If root equals `p`, return root (found target node)
   - If root equals `q`, return root (found target node)

2. **Recursively search subtrees**:
   - Call `lowestCommonAncestor(root.left, p, q)` and store in `left`
   - Call `lowestCommonAncestor(root.right, p, q)` and store in `right`

3. **Determine result based on recursive returns**:
   - If both `left` and `right` are non-null → return root (split point found)
   - If only `left` is non-null → return `left` (both nodes in left subtree)
   - If only `right` is non-null → return `right` (both nodes in right subtree)
   - Can be simplified to: `return left != null ? left : right`

4. **Return the result** (either root, left, or right)

#### Code - Java

```java
public TreeNode lowestCommonAncestor(TreeNode root, TreeNode p, TreeNode q) {
    // Base case: if root is null or matches either node, return root
    if (root == null || root == p || root == q) {
        return root;
    }

    // Recursively search in left and right subtrees
    TreeNode left = lowestCommonAncestor(root.left, p, q);
    TreeNode right = lowestCommonAncestor(root.right, p, q);

    // If both subtrees return non-null, root is the LCA (split point)
    if (left != null && right != null) {
        return root;
    }

    // Otherwise, return whichever subtree found a node (or null if neither)
    return left != null ? left : right;
}
```

**Code Simplification:**
The last two conditions can be combined into a single ternary operator because:
- If `left != null` and `right != null` → we already returned root
- If we reach the final return, at most one of them is non-null
- So we just return whichever is non-null (or null if both are null)


## Binary Tree Upside Down {#lc-upside-down}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 156
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-upside-down/
- **Tags**:
- **Techniques**: [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Flip the tree so that every right child becomes a left sibling.

### Solution - Recursive Rotation

#### Walkthrough

Recurse down the left spine to find the new root. On the way back, rotate pointers: set the left child's left pointer to the current right child, and the left child's right pointer to the current node. Null out the original children to avoid cycles.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java
```
public TreeNode upsideDownBinaryTree(TreeNode root) {
    if (root == null || root.left == null) {
        return root;
    }
    TreeNode newRoot = upsideDownBinaryTree(root.left);
    root.left.left = root.right;
    root.left.right = root;
    root.left = null;
    root.right = null;
    return newRoot;
}
```


## Count Univalue Subtrees {#lc-count-univalue}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 250
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/count-univalue-subtrees/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Count subtrees where every node has the same value.

### Solution - DFS

#### Walkthrough

Traverse bottom-up and return whether each subtree is univalue. A node is univalue if both children are univalue and (when present) match the current node's value. Count each node that satisfies the condition.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java

```
public int countUnivalSubtrees(TreeNode root) {
    int[] count = new int[1];
    dfs(root, count);
    return count[0];
}

private boolean dfs(TreeNode node, int[] count) {
    if (node == null) {
        return true;
    }
    boolean left = dfs(node.left, count);
    boolean right = dfs(node.right, count);
    if (!left || !right) {
        return false;
    }
    if (node.left != null && node.left.val != node.val) {
        return false;
    }
    if (node.right != null && node.right.val != node.val) {
        return false;
    }
    count[0]++;
    return true;
}
```


## Inorder Successor in BST {#lc-inorder-successor}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 285
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/inorder-successor-in-bst/
- **Tags**:
- **Techniques**: [Binary Search Tree](glossary-of-algorithm.html#glossary-bst), Tree

### Description

Find the inorder successor of a given node in a BST.

### Solution - BST Properties

#### Walkthrough

Exploit BST ordering:

- If `p` has a right subtree, successor is the **leftmost node in `p.right`**.
- Otherwise, walk down from root tracking the deepest ancestor whose value is greater than `p.val` (first left turn on the search path).

Iteratively compare `p.val` with `root.val`: when `p.val < root.val`, record `root` as a candidate successor and move left; otherwise move right to search larger values.

#### Analysis

- **Time Complexity**: O(h) where h is tree height
- **Space Complexity**: O(1)

```
public TreeNode inorderSuccessor(TreeNode root, TreeNode p) {
    TreeNode succ = null;
    while (root != null) {
        if (p.val < root.val) {
            succ = root;
            root = root.left;
        } else {
            root = root.right;
        }
    }
    return succ;
}
```


## Binary Tree Longest Consecutive Sequence {#lc-bst-longest-consecutive}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 298
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/binary-tree-longest-consecutive-sequence/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), Tree

### Description

Find the length of the longest path where consecutive nodes differ by +1.

### Solution - DFS Tracking

#### Walkthrough

DFS down the tree while carrying the current consecutive length and the parent value. If the current node extends the sequence (`node.val == parent.val + 1`), increment the length; otherwise reset to 1. Track the maximum over all paths.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java
```
public int longestConsecutive(TreeNode root) {
    return dfs(root, null, 0);
}

private int dfs(TreeNode node, TreeNode parent, int length) {
    if (node == null) {
        return length;
    }
    if (parent != null && node.val == parent.val + 1) {
        length++;
    } else {
        length = 1;
    }
    int left = dfs(node.left, node, length);
    int right = dfs(node.right, node, length);
    return Math.max(length, Math.max(left, right));
}
```


## Find Leaves of Binary Tree {#lc-find-leaves}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 366
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/find-leaves-of-binary-tree/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree

### Description

Collect nodes layer by layer from leaves upward.

### Solution - Height DFS

#### Walkthrough

Compute each node's height measured from the nearest leaf. Leaves have height 0, and a node's height is `1 + max(left, right)`. Use the height as an index into the result list to group nodes by removal round.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java

```
public List<List<Integer>> findLeaves(TreeNode root) {
    List<List<Integer>> result = new ArrayList<>();
    height(root, result);
    return result;
}

private int height(TreeNode node, List<List<Integer>> result) {
    if (node == null) {
        return -1;
    }
    int h = Math.max(height(node.left, result), height(node.right, result)) + 1;
    if (result.size() == h) {
        result.add(new ArrayList<>());
    }
    result.get(h).add(node.val);
    return h;
}
```


## Path Sum III {#lc-path-sum-iii}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 437
- **Difficulty**: Medium
- **URL**: https://leetcode.com/problems/lc-path-sum-iii/
- **Tags**:
- **Techniques**: [Depth First Search](#intro-tree-dfs), Tree, Prefix Sum

### Description

Count the number of paths in a binary tree that sum to a target. Paths may start and end anywhere but must go downward.

### Solution - Prefix Sum DFS

#### Walkthrough

Use DFS while tracking a running prefix sum and a hash map of prefix counts. For each node, check how many previous prefix sums equal `currentSum - target`. Update the map while exploring children and backtrack counts afterward.

#### Analysis

- **Time Complexity**: O(n)
- **Space Complexity**: O(h)

#### Code - Java
```
public int pathSum(TreeNode root, int targetSum) {
    Map<Integer, Integer> prefix = new HashMap<>();
    prefix.put(0, 1);
    return dfs(root, 0, targetSum, prefix);
}

private int dfs(TreeNode node, int current, int target, Map<Integer, Integer> prefix) {
    if (node == null) {
        return 0;
    }

    current += node.val;
    int count = prefix.getOrDefault(current - target, 0);
    prefix.put(current, prefix.getOrDefault(current, 0) + 1);

    count += dfs(node.left, current, target, prefix);
    count += dfs(node.right, current, target, prefix);

    prefix.put(current, prefix.get(current) - 1);
    return count;
}
```


## Subtree of Another Tree {#lc-subtree-of-another-tree}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 572
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/subtree-of-another-tree/
- **Tags**: Blind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), [Recursion](glossary-of-algorithm.html#glossary-recursion), Tree, Binary Tree

### Description

Given the roots of two binary trees `root` and `subRoot`, return `true` if there is a subtree of `root` with the same structure and node values of `subRoot` and `false` otherwise.

A **subtree** of a binary tree `tree` is a tree that consists of a node in `tree` and all of this node's descendants. The tree `tree` could also be considered as a subtree of itself.

### Examples

**Example 1:**
```
Input: root = [3,4,5,1,2], subRoot = [4,1,2]
    3              4
   / \            / \
  4   5          1   2
 / \
1   2

Output: true
Explanation: subRoot has the same structure and node values as a subtree of root.
```

**Example 2:**
```
Input: root = [3,4,5,1,2,null,null,null,null,0], subRoot = [4,1,2]
      3              4
     / \            / \
    4   5          1   2
   / \
  1   2
 /
0

Output: false
Explanation: The subtree with root value 4 in root has an additional node 0.
```

**Example 3:**
```
Input: root = [1,1], subRoot = [1]
Output: true
```

### Constraints

- The number of nodes in the `root` tree is in the range `[1, 2000]`
- The number of nodes in the `subRoot` tree is in the range `[1, 1000]`
- $-10^4 \le \text{root.val} \le 10^4$
- $-10^4 \le \text{subRoot.val} \le 10^4$

### Solution - DFS with Tree Comparison

#### Walkthrough

This solution uses a **two-level DFS approach**:

1. **Outer DFS** (`isSubtree`): Traverse the main tree to find potential matching points
2. **Inner DFS** (`isSame`): Verify if two trees are structurally and value-identical

**Key Insight:** A valid subtree match requires **exact structural and value equality** starting from some node in `root`. The solution explores every node in `root` as a potential subtree root.

**Algorithm Flow:**

1. **Base case handling in `isSubtree`**:
   - If `root` is null, check if `subRoot` is also null (empty tree matches empty tree)
   - This differs from the common approach of returning `false` when `root` is null

2. **Check current node first**:
   - Use `isSame(root, subRoot)` to check if trees match starting from current node
   - If match found, return `true` immediately

3. **Recurse on children**:
   - If current node doesn't match, search left subtree: `isSubtree(root.left, subRoot)`
   - If not found in left, search right subtree: `isSubtree(root.right, subRoot)`
   - Return `true` if found in either subtree

4. **Helper function `isSame`**:
   - Both null → trees are identical (base case)
   - One null, one not → trees differ structurally
   - Values match → recursively verify both left and right subtrees match

**Why This Works:**

The recursive structure ensures we check every possible starting position in `root`. For each position, `isSame` performs a complete structural comparison. The OR operator (`||`) enables early termination once a match is found.

#### Analysis

- **Time Complexity**: O(n × m)
  - n = number of nodes in `root`
  - m = number of nodes in `subRoot`
  - Worst case: we check every node in `root`, and for each node, `isSame` might traverse all m nodes of `subRoot`
  - Example worst case: `root` is a complete binary tree and `subRoot` doesn't exist as a subtree

- **Space Complexity**: O(h₁ + h₂)
  - h₁ = height of `root` tree (recursion stack for `isSubtree`)
  - h₂ = height of `subRoot` tree (recursion stack for `isSame`)
  - In practice, dominated by O(h₁) since `isSame` calls are nested within `isSubtree`
  - Best case (balanced trees): O(log n + log m)
  - Worst case (skewed trees): O(n + m)

**Edge Cases Handled:**

1. Both trees null → `true` (empty subtree exists)
2. Root null, subRoot not null → `false` (can't find non-empty tree in empty tree)
3. SubRoot null, root not null → typically depends on problem definition
4. Single node trees → correctly handled by `isSame`

#### Implementation Steps

1. **Define `isSubtree` main function**:
   - Handle base case: if `root` is null, check if `subRoot` is also null
   - Check current position: call `isSame(root, subRoot)`
   - If match, return `true`
   - Otherwise, recursively search left and right subtrees with OR logic

2. **Define `isSame` helper function**:
   - Base case 1: if both nodes null, return `true`
   - Base case 2: if only one node null, return `false`
   - Recursive case: check value equality AND recursively verify left and right subtrees

3. **Return logic**:
   - `isSubtree` returns `true` if match found anywhere
   - `isSame` returns `true` only if entire structure matches

#### Code - Java

```java
class Solution {
    public boolean isSubtree(TreeNode root, TreeNode subRoot) {
        // Base case: empty root can only match empty subRoot
        if (root == null) {
            return subRoot == null;
        }

        // Check if current node is root of matching subtree
        if (isSame(root, subRoot)) {
            return true;
        }

        // Recursively search in left and right subtrees
        return isSubtree(root.left, subRoot) || isSubtree(root.right, subRoot);
    }

    private boolean isSame(TreeNode root1, TreeNode root2) {
        // Both null: trees are identical
        if (root1 == null && root2 == null) {
            return true;
        }

        // One null, one not: trees are different
        if (root1 == null || root2 == null) {
            return false;
        }

        // Check value equality AND recursive structure match
        return (root1.val == root2.val)
            && isSame(root1.left, root2.left)
            && isSame(root1.right, root2.right);
    }
}
```


## Diameter of Binary Tree {#lc-diameter}

### Problem Metadata

- **Platform**: LeetCode
- **Problem ID**: 543
- **Difficulty**: Easy
- **URL**: https://leetcode.com/problems/diameter-of-binary-tree/
- **Tags**: Grind 75, Blind 75, NeetCode 150
- **Techniques**: [Depth First Search](#intro-tree-dfs), Tree

### Description

Return the length of the longest path between any two nodes.

### Solution - DFS

#### Visual Walkthrough

Consider this tree where the longest path does **not** pass through the root:

```text
        1
       / \
      2   3
     / \
    4   5
   /     \
  6       7
```

**Step-by-Step Trace (Post-order):**

1.  **Leaf Nodes (6, 7, 3)**:
    - `Height = 1`, `Local Diameter = 0 + 0 = 0`
2.  **Node 4**:
    - Left height (from 6) = 1, Right height = 0
    - `Height = 1 + max(1, 0) = 2`
    - `Local Diameter = 1 + 0 = 1`
3.  **Node 5**:
    - Left height = 0, Right height (from 7) = 1
    - `Height = 1 + max(0, 1) = 2`
    - `Local Diameter = 0 + 1 = 1`
4.  **Node 2**:
    - Left height (from 4) = 2, Right height (from 5) = 2
    - `Height = 1 + max(2, 2) = 3`
    - `Local Diameter = 2 + 2 = 4`  <-- **New Global Max**
5.  **Node 1 (Root)**:
    - Left height (from 2) = 3, Right height (from 3) = 1
    - `Height = 1 + max(3, 1) = 4`
    - `Local Diameter = 3 + 1 = 4`

**Final Result**: The maximum diameter encountered was **4** (path: 6 → 4 → 2 → 5 → 7).

#### Walkthrough

The diameter of a binary tree is the longest path between any two nodes. This path may or may not pass through the root. The key insight is that for every node, the longest path that "turns" at that node is the sum of the heights of its left and right subtrees.

**Algorithm:**
1. Use a post-order traversal (DFS) to compute the height of each node's subtrees.
2. For each node, calculate the path length passing through it: `left_height + right_height`.
3. Track the global maximum of these path lengths.
4. Return the height of the current node (`1 + max(left_height, right_height)`) to its parent to facilitate their height calculations.

#### Analysis

- **Time Complexity**: O(n) — each node is visited exactly once.
- **Space Complexity**: O(h) — recursion stack depth depends on tree height.

#### Code - Java

```java
public int diameterOfBinaryTree(TreeNode root) {
    // Use an array to store the global maximum (simulating a pointer/reference)
    int[] diameter = new int[1];
    dfs(root, diameter);
    return diameter[0];
}

private int dfs(TreeNode node, int[] diameter) {
    // Base case: a null node contributes 0 to the path height
    if (node == null) {
        return 0;
    }

    // Recursively find the height of left and right subtrees
    int leftHeight = dfs(node.left, diameter);
    int rightHeight = dfs(node.right, diameter);

    // The diameter passing through this node is the sum of left and right heights
    // Update global maximum if this local path is longer
    diameter[0] = Math.max(diameter[0], leftHeight + rightHeight);

    // Return the height of the current node to the parent
    return 1 + Math.max(leftHeight, rightHeight);
}
```

#### Code - Golang

```go
func diameterOfBinaryTree(root *TreeNode) int {
    // Initialize diameter to 0 (minimum possible value)
    diameter := 0

    // Use a helper function to perform DFS and update diameter
    dfs(root, &diameter)
    return diameter
}

func dfs(node *TreeNode, diameter *int) int {
    // Base case: null nodes have a height of 0
    if node == nil {
        return 0
    }

    // Recursively find the height of left and right subtrees
    left := dfs(node.Left, diameter)
    right := dfs(node.Right, diameter)

    // The diameter passing through this node is left height + right height
    // Update the global maximum if this local path is longer
    *diameter = max(*diameter, left + right)

    // Return the height of the current node (1 + max subtree height)
    return max(left, right) + 1
}
```
