# Searching and Sorting



## Sort a List / Leet Code 148 / Medium

### Description
Sort a linked list in O(n log n) time using constant Auxiliary Space.

### Example
Input:

```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Some caption.', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (A) {4};
\node[list,on chain] (B) {2};
\node[list,on chain] (C) {1};
\node[list,on chain] (D) {3};
\node[on chain,draw,inner sep=6pt] (X) {};
\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);
\draw[*->] let \p1 = (A.two), \p2 = (A.center) in (\x1,\y2) -- (B);
\draw[*->] let \p1 = (B.two), \p2 = (B.center) in (\x1,\y2) -- (C);
\draw[*->] let \p1 = (C.two), \p2 = (C.center) in (\x1,\y2) -- (D);
\draw[*->] let \p1 = (D.two), \p2 = (D.center) in (\x1,\y2) -- (X);
\end{tikzpicture}
```

Output:

```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Some caption.', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (A) {1};
\node[list,on chain] (B) {2};
\node[list,on chain] (C) {3};
\node[list,on chain] (D) {4};
\node[on chain,draw,inner sep=6pt] (X) {};
\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);
\draw[*->] let \p1 = (A.two), \p2 = (A.center) in (\x1,\y2) -- (B);
\draw[*->] let \p1 = (B.two), \p2 = (B.center) in (\x1,\y2) -- (C);
\draw[*->] let \p1 = (C.two), \p2 = (C.center) in (\x1,\y2) -- (D);
\draw[*->] let \p1 = (D.two), \p2 = (D.center) in (\x1,\y2) -- (X);
\end{tikzpicture}
```

### Solution
#### Walkthrough\\
With the idea of Merge sort:

    *  Divide the list in half by finding the middle element of the list. Need to remove the link of middle node
which was previously links to mid - likely the last node in the 1st half list.
    *  Invoke sortList() recurisvely on two sub-lists
    *  Merge and sort against two sub-lists and return head of the list.


#### Analysis\\
In such recursive call to sortList() - O(log(n)) and merge() - O(n), each call creates constant number of temporary
nodes. For all rounds, it will require $O(1)$ of Auxiliary Space. (Unlike time complexity, storage space in recusive
scenario can be reused; and thus, will remain $O(1)$ for all rounds)

#### Algorithm\\
 dnc \@ref(dnc), recursive \@ref(recursive), ms \@ref(ms)

### Java Code
```java
public ListNode sortList(ListNode head) {
    if (head == null || head.next == null) {
        return head;
    }

    ListNode midPrev = null, mid = head, tail = head;

    while (tail != null && tail.next != null) {
        midPrev = mid;

        //mid moves one node at a time
        mid = mid.next;

        //tail moves two node at a time
        tail = tail.next.next;
    }

    //Remove the link of middle node which previously links to mid - likely the last node in the 1st
    //half list
    midPrev.next = null;

    // call sortList() recursively
    ListNode l1 = sortList(head);
    ListNode l2 = sortList(mid);

    return merge(l1, l2);
}

// merge l1 and l2 in sorted order and return the head of sorted list.
private ListNode merge(ListNode l1, ListNode l2) {
    //Two helper nodes, headPrev.next points to the head of merged list
    ListNode current = new ListNode(0), headPrev = current;

    while (l1 != null && l2 != null) {
        if (l1.val < l2.val) {
            current.next = l1;
            l1 = l1.next;
        } else {
            current.next = l2;
            l2 = l2.next;
        }
        current = current.next;
    }

    // l2 is empty with l1 remaining
    if (l1 != null) {
        current.next = l1;
    }

    // l1 is empty with l2 remaining
    if (l2 != null) {
        current.next = l2;
    }

    return headPrev.next;
}
```


## Insertion Sort List / Leet Code 147 / Medium

### Description
Sort a linked list using insertion sort.

Algorithm of Insertion Sort:

    *  Insertion sort iterates, consuming one input element each repetition, and growing a sorted
output list.
    *  At each iteration, insertion sort removes one element from the input data, finds the
location it belongs within the sorted list, and inserts it there.
    *  It repeats until no input elements remain.


### Example
Input:
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (A) {4};
\node[list,on chain] (B) {2};
\node[list,on chain] (C) {1};
\node[list,on chain] (D) {3};
\node[on chain,draw,inner sep=6pt] (X) {};
\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);
\draw[*->] let \p1 = (A.two), \p2 = (A.center) in (\x1,\y2) -- (B);
\draw[*->] let \p1 = (B.two), \p2 = (B.center) in (\x1,\y2) -- (C);
\draw[*->] let \p1 = (C.two), \p2 = (C.center) in (\x1,\y2) -- (D);
\draw[*->] let \p1 = (D.two), \p2 = (D.center) in (\x1,\y2) -- (X);
\end{tikzpicture}

Output:
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (A) {1};
\node[list,on chain] (B) {2};
\node[list,on chain] (C) {3};
\node[list,on chain] (D) {4};
\node[on chain,draw,inner sep=6pt] (X) {};
\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);
\draw[*->] let \p1 = (A.two), \p2 = (A.center) in (\x1,\y2) -- (B);
\draw[*->] let \p1 = (B.two), \p2 = (B.center) in (\x1,\y2) -- (C);
\draw[*->] let \p1 = (C.two), \p2 = (C.center) in (\x1,\y2) -- (D);
\draw[*->] let \p1 = (D.two), \p2 = (D.center) in (\x1,\y2) -- (X);
\end{tikzpicture}

### Solution
#### Walkthrough\\
With Insertion Sort, we could construct a new list by arranging links in place. We need to do the following:

    *  Maintain pointers to node that will be insert-attempted (current) and a previously
sorted nodes.
    *  At each iteration, finds the location it belongs within the sorted list, and inserts between
sorted and sorted.next.
    *  Insert the current node between sorted and sorted.next
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]
\node[list,on chain] (0) {0};
\node[list,on chain] (1) {1};
\node[list,on chain] (2) {5};
\node[list,on chain] (3) {6};
\node[on chain,draw,inner sep=3pt] (X1) {};
\draw (X1.north east) -- (X1.south west);
\draw (X1.north west) -- (X1.south east);
\node[list,on chain] (4)[below= of 2] {2};
\node[on chain,draw,inner sep=3pt] (X2) {};
\draw (X2.north east) -- (X2.south west);
\draw (X2.north west) -- (X2.south east);
\node(headP)[below= of 0] {hp};
\node(sort)[below= of 1] {s};
\node(curr)[below= of 4] {c};

\draw[*->] let \p1 = (0.two), \p2 = (0.center) in (\x1,\y2) -- (1);
\draw[*->, dashed] let \p1 = (1.two), \p2 = (1.center) in (\x1,\y2) -- (2);
\draw[*->] let \p1 = (2.two), \p2 = (2.center) in (\x1,\y2) -- (3);
\draw[*->] let \p1 = (3.two), \p2 = (3.center) in (\x1,\y2) -- (X1);
\draw[*->, dashed] let \p1 = (4.two), \p2 = (4.center) in (\x1,\y2) -- (X2);
\draw[*->] let \p1 = (4.two), \p2 = (4.center) in (\x1,\y2) to[out=60,in=150] (2);
\draw[*->] let \p1 = (1.two), \p2 = (1.center) in (\x1,\y2) to[out=60,in=150] (4);
\draw[*->] (headP) -- (0);
\draw[*->] (sort) -- (1);
\draw[*->] (curr) -- (4);
\end{tikzpicture}
    *  Reset sorted to head of list
    *  It repeats until no input elements remain.


#### Analysis\\
There are two loops, thus time complexity is $O(n^2)$

#### Algorithm\\

### Java Code
```java
public ListNode insertionSortList(ListNode head) {
    if (head == null) {
        return head;
    }

    ListNode headPrev = new ListNode(0); //new starter of the sorted list
    ListNode current = head; //the node will be inserted
    ListNode sorted = headPrev; //insert node between sorted and sorted.next

    while (current != null) {
        ListNode next = current.next;

        //Moving prev until sorted.next.val >= current.val
        while (sorted.next != null && sorted.next.val < current.val) {
            sorted = sorted.next;
        }

        //insert between sorted and sorted.next
        current.next = sorted.next;
        sorted.next = current;

        //move sorted back to head for next round search
        sorted = headPrev;
        current = next;
    }

    return headPrev.next;
}
```


## Selection Sort Array / Firecode / Level 2

### Description

### Example

### Solution
#### Walkthrough\\
Have two nested loops, outer loop move one by one indicating the boundary of unsorted subarray [i...j-1] vs [j ... n (or n-1)],
where the inner loop is to find the minimum element in unsorted array j = [i+1, ... ,n-1]. Swap the found minimum
element with the current element i.
```
[[(1), 5, 6], 2, ...], min_idx = (0)
[[1, 2, 5, 6], ...]
```

#### Analysis\\
There are two loops, thus time complexity is $O(n^2)$

#### Algorithm\\

### Java Code
```java
public static int[] selectionSortArray(int[] arr){
    int n = arr.length;

    // One by one move boundary of unsorted subarray [i...j-1] vs [j ... n]
    for (int i = 0; i < n; i++)
    {
        // Find the minimum element in unsorted array j = [i+1, ... ,n-1]
        int min_idx = i;
        for (int j = i+1; j < n; j++) {
            if (arr[j] < arr[min_idx]) {
                min_idx = j;
            }
        }

        // Swap the found minimum element with the current element
        int temp = arr[min_idx];
        arr[min_idx] = arr[i];
        arr[i] = temp;
    }



    return arr;
}
```


## Quick Sort Array / /

### Description

### Example

### Solution
#### Walkthrough\\
First, we check the indices and continue only if there are still elements to be sorted.  We get the index of the sorted
pivot and use it to recursively call partition() method with the same parameters as the quickSort() method, but with
different indices.

As for the partition() method, this function takes the last element as the pivot. Then, checks each element and swaps
it before the pivot if its value is smaller. By the end of the partitioning, all elements less then the pivot are on
the left of it and all elements greater then the pivot are on the right of it. The pivot is at its final sorted
position and the function returns this position.

#### Analysis\\
The worst case occurs when the partition process always picks greatest or smallest element as pivot and can be $O(n^2)$.
In average case, the time complexity is $O(n \cdot log(n))$

#### Algorithm\\
 qs \@ref(qs)

### Java Code
```java
public void quickSort(int arr[], int begin, int end) {
    if (begin < end) {
        int partitionIndex = partition(arr, begin, end);

        quickSort(arr, begin, partitionIndex-1);
        quickSort(arr, partitionIndex+1, end);
    }
}

//smaller elements go left side of pivot
//larger elements go right side of pivot
private int partition(int arr[], int begin, int end) {
    int pivot = arr[end];

    //index of smaller element
    int i = (begin - 1);

    for (int j = begin; j < end; j++) {
        /*
         * swap if element is smaller than pivot
         */
        if (arr[j] <= pivot) {
            i++;

            int swapTemp = arr[i];
            arr[i] = arr[j];
            arr[j] = swapTemp;
        }
    }

    //swap a[i + 1] and a[end]
    int swapTemp = arr[i + 1];
    arr[i + 1] = arr[end];
    arr[end] = swapTemp;

    return i + 1;
}
```


## Top K elements

### Description
Get k largest elements in an array of size n

### Example

### Solution - Sort array before retrieval
#### Walkthrough\\

#### Analysis\\
\begin{enumearte}
    *  Sort the elements in descending order in $O(n \cdot log(n))$
    *  Get the first k numbers of the sorted array O(k).


#### Algorithm\\

### Java Code  - Sort array before retrieval
```java
public int[] findFirstKElement(int[] arr, int k) {
    int n = arr.length;

    Arrays.sort(arr);

    int [] result = new int[k];
    int index = 0;
    for(int i = n - 1; i >= n - k - 1 && i >= 0; i--) {
        result[index++] = arr[i];
    }

    return result;
}
```

### Solution - Retrieve k element during sort
#### Walkthrough\\

#### Analysis\\
\begin{enumearte}
    *  Modify Bubble Sort to run the outer loop at most k times. $O(n \cdot k)$


#### Algorithm\\
 bs \@ref(bs)

### Java Code  - Retrieve k element during sort
```java
public int[] findFirstKElement(int[] arr, int k) {
    int n = arr.length;

    for (int i = 0; i < n-1 && i < k; i++)  {
        for (int j = 0; j < n-i-1; j++) {
            if (arr[j] > arr[j+1])  {
                // swap arr[j] and arr[j+1]
                int temp = arr[j];
                arr[j] = arr[j+1];
                arr[j+1] = temp;
            }
        }
    }

    int [] result = new int[k];
    int index = 0;
    for(int i = n - 1; i >= n - k - 1 && i >= 0; i--) {
        result[index++] = arr[i];
    }

    return result;
}
```

### Solution - MinHeap
#### Walkthrough\\
Retrieving the top k largest elements equals to retriving smallest element in a k-sized min-heap for k times.The value of
each node is equal to or greater than the value of its parent node and root has the smallest value. Therefore,
we need to maintain a min-heap (PriorityQueue in java) of \textbf{size k} while inserting new element. Below
illustrated how to track a top 2 largest elements from a min heap
```java
3 -> [3*]
2 -> [2*, 3]
1 -> 1, [2*, 3]
5 -> 1, 2, [3*, 5]
6 -> 1, 2, 3, [5*, 6]
4 -> 1, 2, 3, 4, [5*, 6]
```
Finally, return root as it is the smallest element (*) in the heap.

#### Analysis\\
In addition, add() and remove() take $O(log(k))$, since there are at most k element. Thus, total time complexity is
$O(n \cdot log(k))$. Total Auxiliary Space is O(k) due to size of the heap.

#### Algorithm\\
heap \@ref(heap)

### Java Code
```java
public int findFirstKElement(int[] nums, int k) {
    Queue<Integer> heap = new PriorityQueue<>();

    for(int num: nums) {
        helper(num, k);
    }

    int[] result = new int[heap.size()];
    int index = 0;
    while(!heap.isEmpty()) {
        result[index] = heap.remove();
        index++;
    }
    // return the elements from heap in sorted order
    return result;
}

private void helper(int num, int size) {
    heap.add(num);

    if(heap.size() > size) {
        //remove smallest element from min-heap
        heap.remove();
    }
}
```


