# Advanced Data Structure



## Design a LRU Cache / LeetCode 146 / Medium

### Description
Design and implement a data structure for Least Recently Used (LRU) cache. It should support the following
operations: get and put.

get(key) - Get the value (will always be positive) of the key if the key exists in the cache, otherwise return -1.
put(key, value) - Set or insert the value if the key is not already present. When the cache reached its capacity, it
should invalidate the least recently used item before inserting a new item.

The cache is initialized with a positive capacity.
### Example
```java
LRUCache cache = new LRUCache( 2 /* capacity */ );

cache.put(1, 1);
cache.put(2, 2);
cache.get(1);       // returns 1
cache.put(3, 3);    // evicts key 2
cache.get(2);       // returns -1 (not found)
cache.put(4, 4);    // evicts key 1
cache.get(1);       // returns -1 (not found)
cache.get(3);       // returns 3
cache.get(4);       // returns 4
```

### Solution - Naive
#### Walkthrough
Have a internal map to store cache data and doubly linked list to keep track history of data usage. The head
node indicate it was recently used where the tail node represents data that were \textbf{most recently used}.
That is, get(4) would change the cache:
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Some caption.', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (1) {$(3, v_1)$};
\node[list,on chain] (2) {$(2, v_1)$};
\node[list,on chain] (3) {$(4, v_1)$};
\node[on chain,draw,inner sep=3pt] (X) {};
\node[list,on chain] (ph)[below= of 1] {h};

\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);

\draw[*->] let \p1 = (1.two), \p2 = (1.center) in (\x1,\y2) -- (2);
\draw[*->, dashed] let \p1 = (2.two), \p2 = (2.center) in (\x1,\y2) -- (3);
\draw[*->] let \p1 = (2.two), \p2 = (2.center) in (\x1,\y2) [out=120, in=120]to (X);
\draw[*->, dashed] let \p1 = (3.two), \p2 = (3.center) in (\x1,\y2) -- (X);
\draw[*->] let \p1 = (3.two), \p2 = (3.center) in (\x1,\y2) [out=120, in=120]to (1);
\draw[*->, dashed] (ph) -- (1);
\draw[*->] (ph) -- (3);
\end{tikzpicture}
```
In addition, put$(4, v_2)$ would change the cache:
```{r, echo=FALSE, engine='tikz', out.width='90%', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', fig.cap='Some caption.', engine.opts = list(template = "latex/tikz2pdf.tex")}
\begin{tikzpicture}[list/.style={rectangle split, rectangle split parts=2,
draw, rectangle split horizontal}, >=stealth, start chain]

\node[list,on chain] (1) {$(3, v_1)$};
\node[list,on chain] (2) {$(2, v_1)$};
\node[list,on chain] (3) {$(4, v_2)$};
\node[on chain,draw,inner sep=3pt] (X) {};
\node[list,on chain] (ph)[below= of 1] {h};

\draw (X.north east) -- (X.south west);
\draw (X.north west) -- (X.south east);

\draw[*->] let \p1 = (1.two), \p2 = (1.center) in (\x1,\y2) -- (2);
\draw[*->, dashed] let \p1 = (2.two), \p2 = (2.center) in (\x1,\y2) -- (3);
\draw[*->] let \p1 = (2.two), \p2 = (2.center) in (\x1,\y2) [out=120, in=120]to (X);
\draw[*->, dashed] let \p1 = (3.two), \p2 = (3.center) in (\x1,\y2) -- (X);
\draw[*->] let \p1 = (3.two), \p2 = (3.center) in (\x1,\y2) [out=120, in=120]to (1);
\draw[*->, dashed] (ph) -- (1);
\draw[*->] (ph) -- (3);
\end{tikzpicture}
```

#### Analysis
There is a Map used as internal cache and each public method will leverage this internal cache as storage, and thus
each public method cost O(1) overhead. The Auxiliary Space is O(n)

#### Algorithm

### Java Code - Naive
```java
class LRUCache {

    // Doubly Linked List
    class Node {
        int key;

        int data;

        private Node prev;

        private Node next;

        public Node(int key, int data) {
            this.key = key;
            this.data = data;
        }
    }

    private int capacity;

    // internal cache
    private Map<Integer, Node> cache = new HashMap<>();

    private Node head;

    private Node tail;

    public LRUCache(int capacity) {
        this.capacity = capacity;
    }

    public int get(int key) {
        if (cache.containsKey(key)) {
            Node node = cache.get(key);

            //Reset the node to head position
            remove(node);
            setHead(node);

            return node.data;
        }

        return -1;
    }

    public void put(int key, int value) {
        if (cache.containsKey(key)) {
            Node old = cache.get(key);
            old.data = value;

            //Reset the node to head position
            remove(old);
            setHead(old);
        } else {
            // Put new <key, val>
            Node newNode = new Node(key, value);

            if (cache.size() >= capacity) {
                //Cache is FULL. Remove tail
                cache.remove(tail.key); // remove LRU data from cache.
                remove(tail);
                setHead(newNode);
            } else {
                setHead(newNode);
            }

            cache.put(key, newNode);
        }
    }

    private void remove(Node node) {
        if (node == head) {
            // remove a head
            head = node.next;
        } else {
            // remove a non-head
            node.prev.next = node.next;
        }

        if (node == tail) {
            // remove a tail
            tail = node.prev;
        } else {
            // remove a non-head
            node.next.prev = node.prev;
        }
    }

    private void setHead(Node node) {
        node.next = head;
        node.prev = null;

        if (head != null) {
            // insert a new head node
            head.prev = node;
        }
        head = node;

        if (tail == null) {
            // insert first tail node
            tail = head;
        }
    }
}
```

### Solution - LinkedHashMap
#### Walkthrough
An extended LinkedHashMap is to retrieve elements in the order in which they were last accessed.
A special constructor is used to create this type of LinkedHashMap:
new LinkedHashMap(initialCapacity, loadFactor, true) - the true means "accessOrder" instead of the "insertionOrder".

#### Analysis

#### Algorithm

### Java Code - LinkedHashMap
```java
public class LRUCache<K, V> extends LinkedHashMap<K, V> {
    // indicate the capacity of cache.
    private final int maxEntries;
    private static final int DEFAULT_INITIAL_CAPACITY = 16;
    private static final float DEFAULT_LOAD_FACTOR = 0.75f;

    public LRUCache(int initialCapacity, float loadFactor, int maxEntries) {
        super(initialCapacity, loadFactor, true);
        this.maxEntries = maxEntries;
    }

    public LRUCache(int initialCapacity, int maxEntries) {
        this(initialCapacity, DEFAULT_LOAD_FACTOR, maxEntries);
    }

    public LRUCache(int maxEntries) {
        this(DEFAULT_INITIAL_CAPACITY, maxEntries);
    }

    protected boolean removeEldestEntry(Map.Entry<K, V> eldest) {
        return size() > maxEntries;
    }

}
```



## Design a Stack

### Description
Stack is a LIFO data structure.

### Example

    *  push 1: [1]
    *  push 2: [1] $\rightarrow$ [2]
    *  pop returns 2


### Solution - Queue
#### Walkthrough
Since a Queue is FIFO, that is first element will be processed first. Thus, we want to keep
the top element to be inserted first. To achieve that, we could use a temp Stack to reverse
the order after enqueue the added item into the stack first (will be processed first).

#### Analysis

    *  enqueue(item) takes O(n) time.
    *  dequeue() takes O(1) as it is only polling the top element.


#### Algorithm

### Java Code - Queue
```java
class MyQueueBasedStack<T> {
    private Queue<T> value = new LinkedList<>();

    public void push(T item) {
        Queue<T> temp = new LinkedList<>();

        temp.enqueue(item);
        while(!value.isEmpty()) {
            temp.enqueue(value.dequeue());
        }

        value = temp;
    }

    public void pop() {
        return value.poll();
    }

    public T top() {
        return value.peek();
    }

    public boolean empty() {
        return value.isEmpty();
    }

    public int size() {
        return value.size();
    }
}
```



## Design a Queue

### Description
Queue is a FIFO data structure

### Example

    *  enqueue 1: [1]
    *  enqueue 2: [1] $\rightarrow$ [2]
    *  dequeue returns 1


### Solution - Array
#### Walkthrough
With array implementation, there is a need to resize when enqueueing exceeds limits. Thus, in average
case, enqueue is constatnt time but will be O(n) in worst case.

In order to mitigate this cost, we could init the Queue with an immutable large array. Each dequeue will
move an index to the next element. Finally, there could be a background thread to periodically compact
this immutable implementation into a new queue.


    *  enqueue(item) takes O(1) time, but O(n) when resize() is involved.
    %*  enqueue(i, item) takes O(1) time, but O(n) when resize() is involved.
    *  dequeue() takes O(n) time
    %*  dequeue(i) takes O(n) time
    *  search(val) takes O(n) as it traverse all elements.


#### Algorithm

### Java Code - Array
```java
public class MyArrayQueue<T> {

    private T [] array;

    // position of last element
    private int endPos = -1;

    private static final int DEFAULT_INITIAL_CAPACITY = 10;

    public MyArrayQueue() {
        array = (T[]) new Object[DEFAULT_INITIAL_CAPACITY];
    }

    public MyArrayQueue(int initialCapacity) {
        array = (T[]) new Object[initialCapacity];
    }

    public boolean isEmpty() {
        return endPos == -1;
    }

    public int size() {
        return endPos + 1;
    }

    private void resize(int newSize) {
        System.out.println("Resizing from " + array.length + " to " + newSize);
        T[] newArray = Arrays.copyOfRange(array, 0, newSize);
        array = newArray;
    }

    // add the item
    public void enqueue(T item) {
        Objects.requireNonNull(item);

        if( (endPos + 1) >= array.length) {
            resize(2 * array.length);
        }

        endPos++;
        array[endPos] = item;
    }

    // remove first item
    public T dequeue() {
        if(endPos == -1) {
            throw new IllegalStateException();
        }

        // return the first element
        T result = array[0];

        //move the element one backward
        for(int i = 0; i < endPos; i++) {
            array[i] = array[i+1];
        }

        //nullify last position
        array[endPos] = null;
        endPos--;

        return result;
    }

    public boolean search(int val) {
        if(index > endPos) {
            throw new IllegalStateException();
        }


        for(int i = 0; i <= endPos; i++) {
            if(array[i] == val) {
                return true;
            }
        }

        return false;
    }
}
```

### Solution - LinkedList
#### Walkthrough

#### Analysis

    *  enqueue(item) takes O(1) time.
    %*  enqueue(i, item) takes O(n) time to traverse to the previous element.
    *  dequeue() takes O(1)
    %*  dequeue(i) takes O(n) time to traverse to the previous element.
    *  search(val) takes O(n) as it traverses all elements.


#### Algorithm

### Java Code
```java
public class MyLinkedQueue<T> {

    private static class Node<T> {
        Node next;
        T data;
    }

    private Node<T> head;

    private Node<T> tail;

    // position of last element
    private int endPos = -1;

    private static final int DEFAULT_INITIAL_CAPACITY = 10;

    public boolean isEmpty() {
        return endPos == -1;
    }

    public int size() {
        return endPos + 1;
    }

    // add the item
    public void enqueue(T item) {
        Objects.requireNonNull(item);

        endPos++;
        Node newNode = new Node();
        newNode.data = item;

        if(head == null) {
            head = newNode;
        }

        if(tail == null) {
            tail = newNode;
        } else {
            //insert to the tail of the list
            tail.next = newNode;
            tail = newNode;
        }
    }

    // remove and return a random item
    public T dequeue() {
        if(endPos == -1) {
            throw new IllegalStateException();
        }

        // return the first node
        Node dieNode = head;

        //head == tail
        if(head.next == null) {
            head = null;
            tail = null;
        } else {
            head = head.next;
        }
        endPos--;

        return (T) dieNode.data;
    }

    public boolean search(int val) {
        if(index > endPos) {
            throw new IllegalStateException();
        }

        Node cursor = head;
        for(int i = 0; i <= endPos; i++) {
            if(cursor.data == val) {
                return true;
            }
            cursor = cursor.next;
        }

        return false;
    }
}
```

### Solution - Stack
#### Walkthrough
Since a Stack is LIFO, that is last element will be processed first. Thus, we want to keep
the top element to be inserted last. To achieve that, we could use a temp Stack to reverse
the order after pushing the added item into the stack first (will be processed last).

#### Analysis

    *  enqueue(item) takes O(n) time.
    *  dequeue() takes O(1) as it is only popping the top element.


#### Algorithm

### Java Code - Stack
```java
class MyStackBasedQueue<T>() {
    private Stack<T> value = new Stack<>();

    public void enqueue(T item) {
        if(value.isEmpty()) {
            // first element
            value.push(item);
        } else {
            Stack<T> temp = new Stack<>();
            //1. Store previous element in temp
            while(!value.isEmpty()) {
                temp.push(value.pop());
            }

            //2. push item to the back of the stack
            value.push(item);

            //3. reload the previous elements
            while(!temp.isEmpty()) {
                value.push(temp.pop());
            }
        }
    }

    public T dequeue() {
        T result = value.top();
        value.pop();

        return result;
    }

    public int size() {
        return value.size();
    }

    public T peek() {
        return value.top();
    }

    public boolean empty() {
        return value.isEmpty();
    }
}

```



## Design a HashSet / LeetCode 705 / Easy

### Description
Design a HashSet without using any built-in hash table libraries.

To be specific, your design should include these functions:

    *  add(value): Insert a value into the HashSet.
    *  contains(value) : Return whether the value exists in the HashSet or not.
    *  remove(value): Remove a value in the HashSet. If the value does not exist in the HashSet, do nothing.

### Example

### Solution
#### Walkthrough
A HashSet consists of an array of Node where its position is determined by a hash value / hash function. When
there is a collision, we could chain the collided values together using a LinkedList or a Tree. For add()
operation, we need to intensionally skip the duplicated key if the it already exists (from contains(key))


#### Analysis
The performance of each operation heavily depends on effectiveness of hash function. A good hash function would result
in less collision, that is less traversal needed for the chaining list, but likely more auxiliary space required. Best
case for each operation takes O(1). However, In worst case scenario, if it performs all elements on the same chaining
list, all methods can take up to O(n).

#### Algorithm

### Java Code - LinkedList for Chaining
```java
class Node {
    public int key;
    public Node next;
}

class MyHashSet {

    private Node[] list;
    private int size;

    public MyHashSet() {
        this.size = 10000;
        this.list = new Node[this.size];
    }

    public void add(int key) {
        int index = key % this.size;

        if(contains(key)){
            //skip for duplicated entry
        } else {
            //add as head of chaining
            Node oldHead = list[index];

            Node newNode = new Node();
            newNode.key = key;
            newNode.next = oldHead;

            list[index] = newNode;
        }
    }

    public void remove(int key) {
        int index = key % this.size;

        if(contains(key)) {
            Node current = list[index];

            if(current.key == key) {
                // replace head of chaining
                list[index] = current.next;
            } else {
                // replace the middle of chaining
                Node prev = current;

                while(current != null && current.key != key) {
                    prev = current;
                    current = current.next;
                }
                prev.next = current.next;
            }
        } else {
            //skip since not exist
        }
    }

    public boolean contains(int key) {
        int index = key % this.size;

        Node current = list[index];

        while(current != null) {
            if(current.key == key) {
                return true;
            }

            current = current.next;
        }

        return false;
    }
}
```


## Design a HashMap / LeetCode 706 / Easy

### Description
Design a HashMap without using any built-in hash table libraries.

To be specific, your design should include these functions:


    *  put(key, value) : Insert a (key, value) pair into the HashMap. If the value already exists in the HashMap, update the value.
    *  get(key): Returns the value to which the specified key is mapped, or -1 if this map contains no mapping for the key.
    *  remove(key) : Remove the mapping for the value key if this map contains the mapping for the key.

### Example

### Solution
#### Walkthrough
A HashMap consists of an array of Node where its position is determined by a hash value / hash function. When
there is a collision, we could chain the collided values together using a LinkedList or a Tree. One thing different
from HashSet implementation is that we need to update the (key, val) entry if key already exists.

#### Analysis

    *  add() : Best case takes O(1).
    *  get(): Best case takes O(1)
    *  remove(): Best case takes O(1).

However, In worst case scenario, if it performs all elements on the same chaining list, all methods can be O(n).

#### Algorithm

### Java Code - LinkedList for Chaining
```java
class Node {
    public int key;
    public int val;
    public Node next;
}

class MyHashMap {

    private Node[] list;
    private int size;

    /** Initialize your data structure here. */
    public MyHashMap() {
        this.size = 10000;
        this.list = new Node[this.size];
    }

    public void put(int key, int val) {
        Node result = getNode(key);

        if(result != null){
            //if exist, update value
            result.val = val;
        } else {
            Node newNode = new Node();
            newNode.key = key;
            newNode.val = val;

            //add as head of chaining
            int index = key % this.size;
            Node oldHead = list[index];
            newNode.next = oldHead;

            list[index] = newNode;
        }
    }

    public int get(int key) {
        Node result = getNode(key);

        if(result != null) {
            return result.val;
        } else {
            return -1;
        }
    }

    public void remove(int key) {
        int index = key % this.size;

        if(exist(key)) {
            Node head = list[index];

            if(head.key == key) {
                // replace head of chaining
                list[index] = head.next;
            } else {
                // replace the middle of chaining
                Node current = head;
                Node prev = head;

                while(current != null && current.key != key) {
                    prev = current;
                    current = current.next;
                }
                prev.next = current.next;
            }
        } else {
            //skip since not exist
        }
    }

    private boolean exist(int key) {
        return getNode(key) != null;
    }

    private Node getNode(int key) {
        int index = key % this.size;

        Node current = list[index];

        while(current != null) {
            if(current.key == key) {
                return current;
            }

            current = current.next;
        }

        return null;
    }
}
```

