# Problem Index {#problem-index}

This index helps you find problems by LeetCode ID, difficulty, chapter, or platform.

```{r index-setup, include=FALSE}
# Problem database - loaded from CSV file
library(knitr)

# Load all problems from CSV file (105 currently in book)
problems <- read.csv("problems_data.csv", stringsAsFactors = FALSE)

# Ensure techniques column exists for older CSVs
if (!"techniques" %in% names(problems)) {
  problems$techniques <- ""
}

# Normalize missing techniques to empty strings for grepl
problems$techniques[is.na(problems$techniques)] <- ""

# Convert "NA" strings to actual NA values
problems$problem_id[problems$problem_id == "NA"] <- NA
problems$problem_id <- as.numeric(problems$problem_id)

# Add platform priority for sorting (LeetCode=1, HackerRank=2, Firecode=3, Others=4)
platform_order <- c(LeetCode = 1, HackerRank = 2, Firecode = 3, Other = 4, Interview = 4)
problems$platform_priority <- platform_order[problems$platform]
problems$platform_priority[is.na(problems$platform_priority)] <- max(platform_order) + 1

# Add HTML links for clickable problem titles
problems$html_link <- paste0("[", problems$title, "](", problems$anchor, ".html)")

technique_matches <- function(technique) {
  pattern <- paste0("(^|,\\s*)", technique, "(,|$)")
  problems[grepl(pattern, problems$techniques, ignore.case = TRUE), ]
}

# New function for AND queries: finds problems that have ALL specified techniques
technique_matches_and <- function(techniques_vector) {
  result <- problems
  for (technique in techniques_vector) {
    pattern <- paste0("(^|,\\s*)", technique, "(,|$)")
    result <- result[grepl(pattern, result$techniques, ignore.case = TRUE), ]
  }
  result
}

# Function to find problems by technique AND chapter
technique_and_chapter <- function(technique, chapter_name) {
  pattern <- paste0("(^|,\\s*)", technique, "(,|$)")
  problems[grepl(pattern, problems$techniques, ignore.case = TRUE) &
           problems$chapter == chapter_name, ]
}

render_technique_table <- function(technique) {
  matches <- technique_matches(technique)
  if (nrow(matches) == 0) {
    cat("*No problems using this technique yet.*\n")
    return(invisible(NULL))
  }
  matches <- matches[order(matches$platform_priority, matches$problem_id, na.last = TRUE), ]
  matches$display_id <- ifelse(is.na(matches$problem_id), "-", as.character(matches$problem_id))
  matches$problem_link <- paste0("[", matches$title, "](", matches$anchor, ".html)")
  kable(matches[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
}

# New render function for AND queries
render_technique_table_and <- function(techniques_vector) {
  matches <- technique_matches_and(techniques_vector)
  if (nrow(matches) == 0) {
    cat("*No problems using this combination yet.*\n")
    return(invisible(NULL))
  }
  matches <- matches[order(matches$platform_priority, matches$problem_id, na.last = TRUE), ]
  matches$display_id <- ifelse(is.na(matches$problem_id), "-", as.character(matches$problem_id))
  matches$problem_link <- paste0("[", matches$title, "](", matches$anchor, ".html)")
  kable(matches[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
}

# Render function for technique AND chapter queries
render_technique_chapter_table <- function(technique, chapter_name) {
  matches <- technique_and_chapter(technique, chapter_name)
  if (nrow(matches) == 0) {
    cat("*No problems using this combination yet.*\n")
    return(invisible(NULL))
  }
  matches <- matches[order(matches$platform_priority, matches$problem_id, na.last = TRUE), ]
  matches$display_id <- ifelse(is.na(matches$problem_id), "-", as.character(matches$problem_id))
  matches$problem_link <- paste0("[", matches$title, "](", matches$anchor, ".html)")
  kable(matches[, c("platform", "display_id", "problem_link", "difficulty")],
        col.names = c("Platform", "ID", "Problem", "Difficulty"),
        row.names = FALSE)
}
```

---

## Quick Lookup by Problem ID

Browse problems by platform, sorted by problem ID (numeric when present, otherwise by title):

```{r lookup-by-id, echo=FALSE, results='asis'}
platforms <- c("LeetCode", "HackerRank", "Firecode", "Other", "Interview")

for (p in platforms) {
  cat(paste0("### ", p, "\n\n"))
  df <- problems[problems$platform == p, ]
  if (nrow(df) == 0) {
    cat(paste0("*No problems for ", p, " yet.*\n\n"))
    next
  }
  df$problem_id_num <- suppressWarnings(as.numeric(df$problem_id))
  df <- df[order(is.na(df$problem_id_num), df$problem_id_num, df$title, na.last = TRUE), ]
  df$display_id <- ifelse(is.na(df$problem_id_num), "-", as.character(df$problem_id))
  df$problem_link <- paste0("[", df$title, "](", df$anchor, ".html)")
  print(kable(df[, c("display_id", "problem_link", "difficulty", "chapter")],
              col.names = c("ID", "Problem", "Difficulty", "Chapter"),
              row.names = FALSE))
  cat("\n\n")
}
```

---

## Lookup by Difficulty

### Easy Problems

```{r easy-problems, echo=FALSE}
easy <- problems[problems$difficulty == "Easy", ]
# Sort by platform priority, then problem_id
easy <- easy[order(easy$platform_priority, easy$problem_id, na.last = TRUE), ]
easy$display_id <- ifelse(is.na(easy$problem_id), "-", as.character(easy$problem_id))
easy$problem_link <- paste0("[", easy$title, "](", easy$anchor, ".html)")
kable(easy[, c("platform", "display_id", "problem_link", "chapter")],
      col.names = c("Platform", "ID", "Problem", "Chapter"),
      row.names = FALSE)
```

### Medium Problems

```{r medium-problems, echo=FALSE}
medium <- problems[problems$difficulty == "Medium", ]
# Sort by platform priority, then problem_id
medium <- medium[order(medium$platform_priority, medium$problem_id, na.last = TRUE), ]
medium$display_id <- ifelse(is.na(medium$problem_id), "-", as.character(medium$problem_id))
medium$problem_link <- paste0("[", medium$title, "](", medium$anchor, ".html)")
kable(medium[, c("platform", "display_id", "problem_link", "chapter")],
      col.names = c("Platform", "ID", "Problem", "Chapter"),
      row.names = FALSE)
```

### Hard Problems

```{r hard-problems, echo=FALSE}
hard <- problems[problems$difficulty == "Hard", ]
if(nrow(hard) > 0) {
  # Sort by platform priority, then problem_id
  hard <- hard[order(hard$platform_priority, hard$problem_id, na.last = TRUE), ]
  hard$display_id <- ifelse(is.na(hard$problem_id), "-", as.character(hard$problem_id))
  hard$problem_link <- paste0("[", hard$title, "](", hard$anchor, ".html)")
  kable(hard[, c("platform", "display_id", "problem_link", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No hard problems currently in the database.*\n")
}
```

### Unclassified Problems

```{r unclassified-problems, echo=FALSE}
unclassified <- problems[problems$difficulty == "Unclassified", ]
if(nrow(unclassified) > 0) {
  # Sort by platform priority, then problem_id
  unclassified <- unclassified[order(unclassified$platform_priority, unclassified$problem_id, na.last = TRUE), ]
  unclassified$display_id <- ifelse(is.na(unclassified$problem_id), "-", as.character(unclassified$problem_id))
  unclassified$problem_link <- paste0("[", unclassified$title, "](", unclassified$anchor, ".html)")
  kable(unclassified[, c("platform", "display_id", "problem_link", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No unclassified problems currently in the database.*\n")
}
```

---

## Lookup by Chapter

```{r by-chapter, echo=FALSE}
# Create platform priority for sorting (LeetCode=1, HackerRank=2, Firecode=3, Others=4)
platform_order <- c(LeetCode = 1, HackerRank = 2, Firecode = 3, Other = 4, Interview = 4)
problems$platform_priority <- platform_order[problems$platform]
problems$platform_priority[is.na(problems$platform_priority)] <- max(platform_order) + 1

# Sort by chapter, then platform priority, then problem_id (NA last)
by_chapter <- problems[order(problems$chapter_num, problems$platform_priority, problems$problem_id, na.last = TRUE), ]
by_chapter$display_id <- ifelse(is.na(by_chapter$problem_id), "-", as.character(by_chapter$problem_id))
by_chapter$problem_link <- paste0("[", by_chapter$title, "](", by_chapter$anchor, ".html)")
kable(by_chapter[, c("chapter", "problem_link", "platform", "display_id", "difficulty")],
      col.names = c("Chapter", "Problem", "Platform", "Problem ID", "Difficulty"),
      row.names = FALSE)
```

---

## Lookup by Technique

This index helps you find problems by technique. It focuses on uncommon or specialized algorithms and data structures; common techniques like arrays, recursion, and basic iteration are not included since they appear in nearly every problem. Links jump to the detailed explanations in [Common Strategies](#common-strategies) and the [Glossary](#intro).

### Dynamic Programming

#### Tabulation (Bottom-Up)

Problems using [Tabulation](#dp-tabulation):

```{r dp-tabulation-problems, echo=FALSE}
render_technique_table("dp-tabulation")
```

#### Memoization (Top-Down)

Problems using [Memoization](#dp-memoization):

```{r dp-memoization-problems, echo=FALSE}
render_technique_table("dp-memoization")
```

#### Unbounded Knapsack

Problems using the [Unbounded Knapsack](#dp-unbounded-knapsack) pattern:

```{r dp-unbounded-knapsack-problems, echo=FALSE}
render_technique_table("dp-unbounded-knapsack")
```

#### Kadane's Algorithm

Problems using [Kadane's Algorithm](#kadanes-algorithm):

```{r kadanes-algorithm-problems, echo=FALSE}
render_technique_table("kadanes-algorithm")
```

### Backtracking

#### Choose-Explore-Unchoose Pattern

Problems using the [Choose-Explore-Unchoose](#backtracking-choose-explore-unchoose-pattern) pattern:

```{r backtracking-choose-explore-unchoose-problems, echo=FALSE}
render_technique_table("backtracking-choose-explore-unchoose-pattern")
```

#### Pass-by-Value Pattern

Problems using the [Pass-by-Value](#backtracking-pass-by-value-pattern) pattern:

```{r backtracking-pass-by-value-problems, echo=FALSE}
render_technique_table("backtracking-pass-by-value-pattern")
```

#### Mark-Unmark Pattern (Grid or Graph)

Problems using the [Mark-Unmark](#backtracking-mark-unmark-pattern) pattern:

```{r backtracking-mark-unmark-problems, echo=FALSE}
render_technique_table("backtracking-mark-unmark-pattern")
```

### Two Pointer

Problems using the [Two Pointer](#two-pointer) technique:

```{r two-pointer-problems, echo=FALSE}
render_technique_table("two-pointer")
```

### Sliding Window

Problems using the [Sliding Window](#sliding-window) technique:

```{r sliding-window-problems, echo=FALSE}
render_technique_table("sliding-window")
```

### Data Structures

This section indexes problems by the specialized data structures they use.

#### Stack

Problems using the [Stack](#glossary-stack) data structure:

```{r stack-problems, echo=FALSE}
render_technique_table("stack")
```

#### Queue

Problems using the [Queue](#glossary-queue) data structure:

```{r queue-problems, echo=FALSE}
render_technique_table("queue")
```

#### Heap

Problems using the [Heap (Min/Max)](#intro-heap) data structure:

```{r heap-problems, echo=FALSE}
render_technique_table("heap")
```

#### Hash Table

Problems using the [Hash Table](#glossary-hash-table) data structure:

```{r hash-table-problems, echo=FALSE}
render_technique_table("hash-table")
```

#### Binary Search Tree

Problems using the [Binary Search Tree](#glossary-bst) data structure:

```{r bst-problems, echo=FALSE}
render_technique_table("bst")
```

#### Trie

Problems using the [Trie](#glossary-trie) data structure:

```{r trie-problems, echo=FALSE}
render_technique_table("trie")
```

### Searching

This section indexes problems by generic array search algorithms (excluding tree and matrix search).

#### Binary Search

Problems using [Binary Search](#binary-search-strategy):

```{r binary-search-problems, echo=FALSE}
render_technique_table("binary-search")
```

### Matrix Traversal

This section indexes matrix-specific traversal and search problems.

#### Matrix Traversal (General)

Problems using [Matrix Traversal](#intro-matrix-traversal):

```{r matrix-traversal-problems, echo=FALSE}
render_technique_table("matrix-traversal")
```

#### Matrix - Depth-First Search

Problems using [DFS on Matrices](#intro-matrix-dfs):

```{r matrix-dfs-problems, echo=FALSE}
render_technique_chapter_table("dfs", "Matrix")
```

#### Matrix - Breadth-First Search

Problems using [BFS on Matrices](#intro-matrix-bfs):

```{r matrix-bfs-problems, echo=FALSE}
render_technique_chapter_table("bfs", "Matrix")
```

### Tree Traversal

This section indexes tree-specific traversal and search problems.

#### Tree - Depth-First Search

Problems using [DFS on Trees](#intro-tree-dfs):

```{r tree-dfs-problems, echo=FALSE}
render_technique_chapter_table("dfs", "Tree")
```

#### Tree - Breadth-First Search

Problems using [BFS on Trees](#intro-tree-bfs):

```{r tree-bfs-problems, echo=FALSE}
render_technique_chapter_table("bfs", "Tree")
```

### Graph Traversal

This section indexes problems by specific graph traversal techniques, organized by graph type and algorithm.

#### Undirected Graph - Depth First Search

Problems using [Undirected Graph DFS](#undirected-graph-dfs):

```{r undirected-graph-dfs-problems, echo=FALSE}
render_technique_table("undirected-graph-dfs")
```

#### Undirected Graph - Breadth First Search

Problems using [Undirected Graph BFS](#undirected-graph-bfs):

```{r undirected-graph-bfs-problems, echo=FALSE}
render_technique_table("undirected-graph-bfs")
```

#### Directed Acyclic Graph (DAG) - Depth First Search

Problems using [DAG DFS](#dag-dfs) for topological sorting:

```{r dag-dfs-problems, echo=FALSE}
render_technique_table("dag-dfs")
```

#### Directed Acyclic Graph (DAG) - Breadth First Search

Problems using [DAG BFS](#dag-bfs) (Kahn's Algorithm) for topological sorting:

```{r dag-bfs-problems, echo=FALSE}
render_technique_table("dag-bfs")
```

#### Cycle Detection in Undirected Graph

Problems using [Cycle Detection in Undirected Graph](#undirected-graph-cycle-detection):

```{r undirected-graph-cycle-detection-problems, echo=FALSE}
render_technique_table("undirected-graph-cycle-detection")
```

#### Cycle Detection in DAG

Problems using [Cycle Detection in DAG](#dag-cycle-detection):

```{r dag-cycle-detection-problems, echo=FALSE}
render_technique_table("dag-cycle-detection")
```

### Sorting

This section indexes problems by sorting algorithms and techniques.

#### Bubble Sort

Problems using [Bubble Sort](#sorting-bubble-sort):

```{r bubble-sort-problems, echo=FALSE}
render_technique_table("sorting-bubble-sort")
```

#### Insertion Sort

Problems using [Insertion Sort](#sorting-insertion-sort):

```{r insertion-sort-problems, echo=FALSE}
render_technique_table("sorting-insertion-sort")
```

#### Merge Sort

Problems using [Merge Sort](#sorting-merge-sort):

```{r merge-sort-problems, echo=FALSE}
render_technique_table("sorting-merge-sort")
```

#### Quick Sort

Problems using [Quick Sort](#sorting-quick-sort):

```{r quick-sort-problems, echo=FALSE}
render_technique_table("sorting-quick-sort")
```

#### Topological Sort

Problems using [Topological Sort](#sorting-topological-sort):

```{r topological-sort-problems, echo=FALSE}
render_technique_table("sorting-topological-sort")
```

### Statistical Sampling

#### Rejection Sampling

Problems using [Rejection Sampling](#glossary-rejection-sampling):

```{r rejection-sampling-problems, echo=FALSE}
render_technique_table("rejection-sampling")
```

#### Reservoir Sampling

Problems using [Reservoir Sampling](#glossary-reservoir-sampling):

```{r reservoir-sampling-problems, echo=FALSE}
render_technique_table("reservoir-sampling")
```

### Game Theory & Optimization

#### MinMax

Problems using the [MinMax](#glossary-minimax) algorithm:

```{r minimax-problems, echo=FALSE}
render_technique_table("minimax")
```

---

## Lookup by Study Plan

### Blind 75

```{r blind-75, echo=FALSE}
blind75 <- problems[grepl("Blind 75", problems$tags, ignore.case = TRUE), ]
cat(sprintf("*Progress: %d / 75 problems.*\n\n", nrow(blind75)))
if(nrow(blind75) > 0) {
  # Sort by platform priority, then problem_id
  blind75 <- blind75[order(blind75$platform_priority, blind75$problem_id, na.last = TRUE), ]
  blind75$display_id <- ifelse(is.na(blind75$problem_id), "-", as.character(blind75$problem_id))
  blind75$problem_link <- paste0("[", blind75$title, "](", blind75$anchor, ".html)")
  kable(blind75[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No problems tagged as Blind 75 yet. Tags will be populated by Codex.*\n")
}
```

### NeetCode 150

```{r neetcode-150, echo=FALSE}
neetcode <- problems[grepl("NeetCode 150", problems$tags, ignore.case = TRUE), ]
cat(sprintf("*Progress: %d / 150 problems.*\n\n", nrow(neetcode)))
if(nrow(neetcode) > 0) {
  # Sort by platform priority, then problem_id
  neetcode <- neetcode[order(neetcode$platform_priority, neetcode$problem_id, na.last = TRUE), ]
  neetcode$display_id <- ifelse(is.na(neetcode$problem_id), "-", as.character(neetcode$problem_id))
  neetcode$problem_link <- paste0("[", neetcode$title, "](", neetcode$anchor, ".html)")
  kable(neetcode[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No problems tagged as NeetCode 150 yet. Tags will be populated by Codex.*\n")
}
```

### Grind 75

```{r grind-75, echo=FALSE}
grind75 <- problems[grepl("Grind 75", problems$tags, ignore.case = TRUE), ]
cat(sprintf("*Progress: %d / 75 problems.*\n\n", nrow(grind75)))
if(nrow(grind75) > 0) {
  # Sort by platform priority, then problem_id
  grind75 <- grind75[order(grind75$platform_priority, grind75$problem_id, na.last = TRUE), ]
  grind75$display_id <- ifelse(is.na(grind75$problem_id), "-", as.character(grind75$problem_id))
  grind75$problem_link <- paste0("[", grind75$title, "](", grind75$anchor, ".html)")
  kable(grind75[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No problems tagged as Grind 75 yet. Tags will be populated by Codex.*\n")
}
```

---

## Lookup by Company

```{r company-tags, echo=FALSE}
# Extract all unique company tags from the tags column
# Common companies to check for
companies <- c("Google", "Amazon", "Meta", "Microsoft", "Apple", "Bloomberg", "Adobe", "Uber", "Netflix", "Facebook")

# Function to display problems for a company
show_company_problems <- function(company_name) {
  company_problems <- problems[grepl(company_name, problems$tags, ignore.case = TRUE), ]
  if(nrow(company_problems) > 0) {
    # Sort by platform priority, then problem_id
    company_problems <- company_problems[order(company_problems$platform_priority, company_problems$problem_id, na.last = TRUE), ]
    company_problems$display_id <- ifelse(is.na(company_problems$problem_id), "-", as.character(company_problems$problem_id))
    company_problems$problem_link <- paste0("[", company_problems$title, "](", company_problems$anchor, ".html)")
    cat(paste0("\n### ", company_name, " (", nrow(company_problems), " problems)\n\n"))
    print(kable(company_problems[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
          col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
          row.names = FALSE))
  }
}

# Check if any company tags exist
has_company_tags <- any(sapply(companies, function(c) any(grepl(c, problems$tags, ignore.case = TRUE))))

if(has_company_tags) {
  for(company in companies) {
    show_company_problems(company)
  }
} else {
  cat("*No company tags found yet. Tags will be populated by Codex.*\n\nOnce populated, this section will show problems grouped by company (Google, Amazon, Meta, etc.).\n")
}
```

---

## How to Update This Index

When you add a new problem to the book:

1. Open `problems_data.csv` in a spreadsheet or text editor
2. Add a new row with the required fields: title, platform, problem_id, difficulty, chapter, chapter_num, anchor, tags, techniques
3. Use comma-separated values in the tags and techniques columns, wrapping the entire value in quotes when it contains commas
4. Rebuild the book with `sh build.sh`
5. All lookup tables will automatically update

### Example CSV Entry

```csv
Your New Problem,LeetCode,9999,Medium,Array,02,new-problem,"Google, Amazon, Top 100 Liked, Blind 75","dp-tabulation, hash-table"
```

### Tags Format

The `tags` field supports comma-separated values for:

- **Company names**: Google, Amazon, Meta, Microsoft, Apple, Bloomberg, Adobe, Uber, Netflix
- **Study plans**: Top 100 Liked, Blind 75, NeetCode 150, Grind 75

Use quotes around the tags value if it contains commas.

### Techniques Format

The `techniques` field supports comma-separated technique anchor IDs for:

- **Dynamic Programming**: `dp-memoization`, `dp-tabulation`, `dp-unbounded-knapsack`, `kadanes-algorithm`
- **Backtracking**: `backtracking-choose-explore-unchoose-pattern`, `backtracking-pass-by-value-pattern`, `backtracking-mark-unmark-pattern`
- **Two Pointer**: `two-pointer`, `sliding-window`
- **Data Structures**: `stack`, `queue`, `heap`, `hash-table`, `bst`, `trie`, `linked-list`, `array`
- **Searching**: `binary-search`
- **Matrix Traversal**: `matrix-traversal`, `dfs` (for matrices), `bfs` (for matrices)
- **Tree Traversal**: `dfs` (for trees), `bfs` (for trees)
- **Graph Traversal**: `undirected-graph-dfs`, `undirected-graph-bfs`, `dag-dfs`, `dag-bfs`, `undirected-graph-cycle-detection`, `dag-cycle-detection`
- **Sorting**: `sorting-bubble-sort`, `sorting-insertion-sort`, `sorting-merge-sort`, `sorting-quick-sort`, `sorting-topological-sort`
- **Statistical Sampling**: `rejection-sampling`, `reservoir-sampling`
- **Paradigms**: `greedy`, `dnc`, `recursion`
- **Matrix Operations**: `matrix-manipulation`
- **Advanced**: `bitwise`, `simulation`, `probability`, `minimax`

Use quotes around the techniques value if it contains commas.

**Note**: When using `dfs` or `bfs` techniques, the problem will be categorized by chapter:

- Tree chapter results in Tree DFS or Tree BFS categorization
- Matrix chapter results in Matrix DFS or Matrix BFS categorization
- Graph chapter results in Graph-specific traversal patterns

---

## Summary Statistics

```{r stats, echo=FALSE}
cat("**Total Problems:**", nrow(problems), "\n\n")
cat("**By Difficulty:**\n")
cat("- Easy:", sum(problems$difficulty == "Easy"), "\n")
cat("- Medium:", sum(problems$difficulty == "Medium"), "\n")
cat("- Hard:", sum(problems$difficulty == "Hard"), "\n")
cat("- Unclassified:", sum(problems$difficulty == "Unclassified"), "\n\n")
cat("**By Platform:**\n")
for(p in unique(problems$platform)) {
  cat("-", p, ":", sum(problems$platform == p), "\n")
}
```

---

*This index is automatically generated from `problems_data.csv`. All problem titles are clickable links to the generated HTML documentation in `docs/`.*
