# Problem Index {#problem-index}

This index helps you find problems by LeetCode ID, difficulty, chapter, or platform.

```{r index-setup, include=FALSE}
# Problem database - loaded from CSV file
library(knitr)

# Load all problems from CSV file (105 currently in book)
problems <- read.csv("problems_data.csv", stringsAsFactors = FALSE)

# Convert "NA" strings to actual NA values
problems$problem_id[problems$problem_id == "NA"] <- NA
problems$problem_id <- as.numeric(problems$problem_id)

# Add platform priority for sorting (LeetCode=1, HackerRank=2, Firecode=3, Others=4)
platform_order <- c(LeetCode = 1, HackerRank = 2, Firecode = 3, Other = 4, Interview = 4)
problems$platform_priority <- platform_order[problems$platform]
problems$platform_priority[is.na(problems$platform_priority)] <- max(platform_order) + 1

# Add HTML links for clickable problem titles
problems$html_link <- paste0("[", problems$title, "](", problems$anchor, ".html)")
```

---

## Quick Lookup by Problem ID

Browse problems by platform, sorted by problem ID (numeric when present, otherwise by title):

```{r lookup-by-id, echo=FALSE, results='asis'}
platforms <- c("LeetCode", "HackerRank", "Firecode", "Other", "Interview")

for (p in platforms) {
  cat(paste0("### ", p, "\n\n"))
  df <- problems[problems$platform == p, ]
  if (nrow(df) == 0) {
    cat(paste0("*No problems for ", p, " yet.*\n\n"))
    next
  }
  df$problem_id_num <- suppressWarnings(as.numeric(df$problem_id))
  df <- df[order(is.na(df$problem_id_num), df$problem_id_num, df$title, na.last = TRUE), ]
  df$display_id <- ifelse(is.na(df$problem_id_num), "-", as.character(df$problem_id))
  df$problem_link <- paste0("[", df$title, "](", df$anchor, ".html)")
  print(kable(df[, c("display_id", "problem_link", "difficulty", "chapter")],
              col.names = c("ID", "Problem", "Difficulty", "Chapter"),
              row.names = FALSE))
  cat("\n\n")
}
```

---

## Lookup by Difficulty

### Easy Problems

```{r easy-problems, echo=FALSE}
easy <- problems[problems$difficulty == "Easy", ]
# Sort by platform priority, then problem_id
easy <- easy[order(easy$platform_priority, easy$problem_id, na.last = TRUE), ]
easy$display_id <- ifelse(is.na(easy$problem_id), "-", as.character(easy$problem_id))
easy$problem_link <- paste0("[", easy$title, "](", easy$anchor, ".html)")
kable(easy[, c("platform", "display_id", "problem_link", "chapter")],
      col.names = c("Platform", "ID", "Problem", "Chapter"),
      row.names = FALSE)
```

### Medium Problems

```{r medium-problems, echo=FALSE}
medium <- problems[problems$difficulty == "Medium", ]
# Sort by platform priority, then problem_id
medium <- medium[order(medium$platform_priority, medium$problem_id, na.last = TRUE), ]
medium$display_id <- ifelse(is.na(medium$problem_id), "-", as.character(medium$problem_id))
medium$problem_link <- paste0("[", medium$title, "](", medium$anchor, ".html)")
kable(medium[, c("platform", "display_id", "problem_link", "chapter")],
      col.names = c("Platform", "ID", "Problem", "Chapter"),
      row.names = FALSE)
```

### Hard Problems

```{r hard-problems, echo=FALSE}
hard <- problems[problems$difficulty == "Hard", ]
if(nrow(hard) > 0) {
  # Sort by platform priority, then problem_id
  hard <- hard[order(hard$platform_priority, hard$problem_id, na.last = TRUE), ]
  hard$display_id <- ifelse(is.na(hard$problem_id), "-", as.character(hard$problem_id))
  hard$problem_link <- paste0("[", hard$title, "](", hard$anchor, ".html)")
  kable(hard[, c("platform", "display_id", "problem_link", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No hard problems currently in the database.*\n")
}
```

### Unclassified Problems

```{r unclassified-problems, echo=FALSE}
unclassified <- problems[problems$difficulty == "Unclassified", ]
if(nrow(unclassified) > 0) {
  # Sort by platform priority, then problem_id
  unclassified <- unclassified[order(unclassified$platform_priority, unclassified$problem_id, na.last = TRUE), ]
  unclassified$display_id <- ifelse(is.na(unclassified$problem_id), "-", as.character(unclassified$problem_id))
  unclassified$problem_link <- paste0("[", unclassified$title, "](", unclassified$anchor, ".html)")
  kable(unclassified[, c("platform", "display_id", "problem_link", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No unclassified problems currently in the database.*\n")
}
```

---

## Lookup by Chapter

```{r by-chapter, echo=FALSE}
# Create platform priority for sorting (LeetCode=1, HackerRank=2, Firecode=3, Others=4)
platform_order <- c(LeetCode = 1, HackerRank = 2, Firecode = 3, Other = 4, Interview = 4)
problems$platform_priority <- platform_order[problems$platform]
problems$platform_priority[is.na(problems$platform_priority)] <- max(platform_order) + 1

# Sort by chapter, then platform priority, then problem_id (NA last)
by_chapter <- problems[order(problems$chapter_num, problems$platform_priority, problems$problem_id, na.last = TRUE), ]
by_chapter$display_id <- ifelse(is.na(by_chapter$problem_id), "-", as.character(by_chapter$problem_id))
by_chapter$problem_link <- paste0("[", by_chapter$title, "](", by_chapter$anchor, ".html)")
kable(by_chapter[, c("chapter", "problem_link", "platform", "display_id", "difficulty")],
      col.names = c("Chapter", "Problem", "Platform", "Problem ID", "Difficulty"),
      row.names = FALSE)
```

---

## Lookup by Study Plan

### Blind 75

```{r blind-75, echo=FALSE}
blind75 <- problems[grepl("Blind 75", problems$tags, ignore.case = TRUE), ]
cat(sprintf("*Progress: %d / 75 problems.*\n\n", nrow(blind75)))
if(nrow(blind75) > 0) {
  # Sort by platform priority, then problem_id
  blind75 <- blind75[order(blind75$platform_priority, blind75$problem_id, na.last = TRUE), ]
  blind75$display_id <- ifelse(is.na(blind75$problem_id), "-", as.character(blind75$problem_id))
  blind75$problem_link <- paste0("[", blind75$title, "](", blind75$anchor, ".html)")
  kable(blind75[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No problems tagged as Blind 75 yet. Tags will be populated by Codex.*\n")
}
```

### NeetCode 150

```{r neetcode-150, echo=FALSE}
neetcode <- problems[grepl("NeetCode 150", problems$tags, ignore.case = TRUE), ]
cat(sprintf("*Progress: %d / 150 problems.*\n\n", nrow(neetcode)))
if(nrow(neetcode) > 0) {
  # Sort by platform priority, then problem_id
  neetcode <- neetcode[order(neetcode$platform_priority, neetcode$problem_id, na.last = TRUE), ]
  neetcode$display_id <- ifelse(is.na(neetcode$problem_id), "-", as.character(neetcode$problem_id))
  neetcode$problem_link <- paste0("[", neetcode$title, "](", neetcode$anchor, ".html)")
  kable(neetcode[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
        col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
        row.names = FALSE)
} else {
  cat("*No problems tagged as NeetCode 150 yet. Tags will be populated by Codex.*\n")
}
```

---

## Lookup by Company

```{r company-tags, echo=FALSE}
# Extract all unique company tags from the tags column
# Common companies to check for
companies <- c("Google", "Amazon", "Meta", "Microsoft", "Apple", "Bloomberg", "Adobe", "Uber", "Netflix", "Facebook")

# Function to display problems for a company
show_company_problems <- function(company_name) {
  company_problems <- problems[grepl(company_name, problems$tags, ignore.case = TRUE), ]
  if(nrow(company_problems) > 0) {
    # Sort by platform priority, then problem_id
    company_problems <- company_problems[order(company_problems$platform_priority, company_problems$problem_id, na.last = TRUE), ]
    company_problems$display_id <- ifelse(is.na(company_problems$problem_id), "-", as.character(company_problems$problem_id))
    company_problems$problem_link <- paste0("[", company_problems$title, "](", company_problems$anchor, ".html)")
    cat(paste0("\n### ", company_name, " (", nrow(company_problems), " problems)\n\n"))
    print(kable(company_problems[, c("platform", "display_id", "problem_link", "difficulty", "chapter")],
          col.names = c("Platform", "ID", "Problem", "Difficulty", "Chapter"),
          row.names = FALSE))
  }
}

# Check if any company tags exist
has_company_tags <- any(sapply(companies, function(c) any(grepl(c, problems$tags, ignore.case = TRUE))))

if(has_company_tags) {
  for(company in companies) {
    show_company_problems(company)
  }
} else {
  cat("*No company tags found yet. Tags will be populated by Codex.*\n\nOnce populated, this section will show problems grouped by company (Google, Amazon, Meta, etc.).\n")
}
```

---

## How to Update This Index

When you add a new problem to the book:

1. Open [problems_data.csv](problems_data.csv) in a spreadsheet or text editor
2. Add a new row with: `title,platform,problem_id,difficulty,chapter,chapter_num,anchor,tags`
3. Rebuild the book with `sh build.sh`
4. All lookup tables will automatically update!

### Example CSV Entry

```
Your New Problem,LeetCode,9999,Medium,Array,02,new-problem,"Google, Amazon, Top 100 Liked, Blind 75"
```

### Tags Format

The `tags` field supports comma-separated values for:
- **Company names**: Google, Amazon, Meta, Microsoft, Apple, Bloomberg, Adobe, Uber, Netflix
- **Study plans**: Top 100 Liked, Blind 75, NeetCode 150

Use quotes around the tags value if it contains commas.

---

## Summary Statistics

```{r stats, echo=FALSE}
cat("**Total Problems:**", nrow(problems), "\n\n")
cat("**By Difficulty:**\n")
cat("- Easy:", sum(problems$difficulty == "Easy"), "\n")
cat("- Medium:", sum(problems$difficulty == "Medium"), "\n")
cat("- Hard:", sum(problems$difficulty == "Hard"), "\n")
cat("- Unclassified:", sum(problems$difficulty == "Unclassified"), "\n\n")
cat("**By Platform:**\n")
for(p in unique(problems$platform)) {
  cat("-", p, ":", sum(problems$platform == p), "\n")
}
```

---

*This index is automatically generated from `problems_data.csv`. All problem titles are clickable links to the generated HTML documentation in `docs/`.*
