# Problem Index {#problem-index}

This index helps you find problems by LeetCode ID, difficulty, chapter, or platform.

```{r index-setup, include=FALSE}
# Problem database - loaded from CSV file
library(knitr)

# Load all problems from CSV file (105 currently in book)
problems <- read.csv("problems_data.csv", stringsAsFactors = FALSE)

# Convert "NA" strings to actual NA values
problems$problem_id[problems$problem_id == "NA"] <- NA
problems$problem_id <- as.numeric(problems$problem_id)

# Add HTML links for clickable problem titles
problems$html_link <- paste0("[", problems$title, "](", problems$anchor, ".html)")
```

---

## Quick Lookup by LeetCode ID

Find any LeetCode problem by its problem number:

```{r lookup-by-id, echo=FALSE}
leetcode_only <- problems[problems$platform == "LeetCode" & !is.na(problems$problem_id), ]
leetcode_sorted <- leetcode_only[order(leetcode_only$problem_id), ]

# Add clickable links
leetcode_sorted$problem_link <- paste0("[", leetcode_sorted$title, "](", leetcode_sorted$anchor, ".html)")

kable(leetcode_sorted[, c("problem_id", "problem_link", "difficulty", "chapter")],
      col.names = c("ID", "Problem", "Difficulty", "Chapter"),
      row.names = FALSE)
```

---

## Lookup by Difficulty

### Easy Problems

```{r easy-problems, echo=FALSE}
easy <- problems[problems$difficulty == "Easy", ]
easy$display_id <- ifelse(is.na(easy$problem_id), "-", as.character(easy$problem_id))
easy$problem_link <- paste0("[", easy$title, "](", easy$anchor, ".html)")
kable(easy[, c("display_id", "problem_link", "chapter", "platform")],
      col.names = c("ID", "Problem", "Chapter", "Platform"),
      row.names = FALSE)
```

### Medium Problems

```{r medium-problems, echo=FALSE}
medium <- problems[problems$difficulty == "Medium", ]
medium$display_id <- ifelse(is.na(medium$problem_id), "-", as.character(medium$problem_id))
medium$problem_link <- paste0("[", medium$title, "](", medium$anchor, ".html)")
kable(medium[, c("display_id", "problem_link", "chapter", "platform")],
      col.names = c("ID", "Problem", "Chapter", "Platform"),
      row.names = FALSE)
```

### Hard Problems

```{r hard-problems, echo=FALSE}
hard <- problems[problems$difficulty == "Hard", ]
if(nrow(hard) > 0) {
  hard$display_id <- ifelse(is.na(hard$problem_id), "-", as.character(hard$problem_id))
  hard$problem_link <- paste0("[", hard$title, "](", hard$anchor, ".html)")
  kable(hard[, c("display_id", "problem_link", "chapter", "platform")],
        col.names = c("ID", "Problem", "Chapter", "Platform"),
        row.names = FALSE)
} else {
  cat("*No hard problems currently in the database.*\n")
}
```

### Unclassified Problems

```{r unclassified-problems, echo=FALSE}
unclassified <- problems[problems$difficulty == "Unclassified", ]
if(nrow(unclassified) > 0) {
  unclassified$display_id <- ifelse(is.na(unclassified$problem_id), "-", as.character(unclassified$problem_id))
  unclassified$problem_link <- paste0("[", unclassified$title, "](", unclassified$anchor, ".html)")
  kable(unclassified[, c("display_id", "problem_link", "chapter", "platform")],
        col.names = c("ID", "Problem", "Chapter", "Platform"),
        row.names = FALSE)
} else {
  cat("*No unclassified problems currently in the database.*\n")
}
```

---

## Lookup by Chapter

```{r by-chapter, echo=FALSE}
by_chapter <- problems[order(problems$chapter_num, problems$problem_id, na.last = TRUE), ]
by_chapter$display_id <- ifelse(is.na(by_chapter$problem_id), "-", as.character(by_chapter$problem_id))
by_chapter$problem_link <- paste0("[", by_chapter$title, "](", by_chapter$anchor, ".html)")
kable(by_chapter[, c("chapter", "problem_link", "display_id", "difficulty")],
      col.names = c("Chapter", "Problem", "LeetCode ID", "Difficulty"),
      row.names = FALSE)
```

---

## How to Update This Index

When you add a new problem to the book:

1. Open [problems_data.csv](problems_data.csv) in a spreadsheet or text editor
2. Add a new row with: `title,platform,problem_id,difficulty,chapter,chapter_num,anchor`
3. Rebuild the book with `sh build.sh`
4. All lookup tables will automatically update!

### Example CSV Entry

```
Your New Problem,LeetCode,9999,Medium,Array,02,new-problem
```

---

## Summary Statistics

```{r stats, echo=FALSE}
cat("**Total Problems:**", nrow(problems), "\n\n")
cat("**By Difficulty:**\n")
cat("- Easy:", sum(problems$difficulty == "Easy"), "\n")
cat("- Medium:", sum(problems$difficulty == "Medium"), "\n")
cat("- Hard:", sum(problems$difficulty == "Hard"), "\n")
cat("- Unclassified:", sum(problems$difficulty == "Unclassified"), "\n\n")
cat("**By Platform:**\n")
for(p in unique(problems$platform)) {
  cat("-", p, ":", sum(problems$platform == p), "\n")
}
```

---

*This index is automatically generated from `problems_data.csv`. All problem titles are clickable links to the generated HTML documentation in `docs/`.*
